### 2021-08-20 documentos/notas/notas_carlos.tex

\section{Kraus}

In order to study the Kraus operators corresponding to PCE's, it is enough to study 
the ones corresponding to generators, that is, the ones that leave 
have of the components invariant.

As a historical fact, the Kraus operators corresponding to one single qubit PCEs
are 
\begin{equation}
\{\1/\sqrt2, \sigma_i/\sqrt2\}
\label{eq:kraus:PCE}
\end{equation}
(corresponding to the operation that leaves the component $\sigma_i$ invariant), 
see Eq. 8.95 of Chuang's book.  Visual inspection of the Kraus operators 
of two qubit PCEG (PCE generators) lead to the ansatz that 
the appropriate Kraus operators are 
\begin{equation}
\left\{\1/\sqrt2, \left(\bigotimes_k \sigma_{i_k}\right)/\sqrt2\right\}. 
\label{}
\end{equation}

We shall first show that the aforementioned Kraus operators indeed produce a PCE. 
Notice that 
\begin{equation}
\sigma_i \sigma_j \sigma_i =a^i_{j} \sigma_j,
\label{}
\end{equation}
with $a$ the matrix that appears in 
\eqref{eq:eigvals-PCE}. Next, consider the action of a channel with Kraus representation 
\eqref{eq:kraus:PCE} on an $n$ qubit system with Bloch sphere representation \eqref{eq:nqubit:bloch},
\begin{align}
M[\rho] &= 
\frac{\1}{\sqrt{2}} \rho\frac{\1}{\sqrt{2}} 
 + 
\frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \rho
\frac{\otimes_k \sigma_{i_k}}{\sqrt{2}} \\
&=\frac12 \left(\rho + \bigotimes_k \sigma_{i_k} \rho \bigotimes_k \sigma_{i_k} \right)\\
&=\frac12 \rho
+ \frac12\sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \bigotimes_k  \sigma_{i_k} \sigma_{j_k}  \sigma_{i_k} \\
&=
\sum_{j_1, \ldots,j_n} r_{j_1 \ldots j_n} \frac{1+\prod_k a^{i_k}_{j_k}}{2} \bigotimes_k  \sigma_{j_k} . 
\end{align}
However, $(1+\prod_k a^{i_k}_{j_k})/2$ is 0 or 1, proving the channel is a PCE.

We still have to show that except for $i_1=\cdots=i_n=0$, half of the numbers 
$(1+\prod_k a^{i_k}_{j_k})/2=1$, or equivalently, half of $\prod_k a^{i_k}_{j_k}=1$. 
Consider a particular $l$ such that $i_l \ne 0$. This will make half of 
the $\prod_k a^{i_k}_{j_k}$ 1 and half $-1$ (for fixed $l$). This also implies 
that half of the correlations of the density matrix are erased and half preserved 
so the channel is a PCEG.

Notice also that different choices of sets $\{i_k \}$ lead to different channels. This
can be seen noticing that if two channels were the same, this would imply that
the matrix representation of the corresponding superoperator of 
$\bigotimes_k \sigma_{i_k} \cdot \bigotimes_k \sigma_{i_k} $ would have to be
the same which is clearly false.

Two remarks. (i) The notation that we develop with Jose Alfredo coincides with the 
generator. This notation is based on the action on each individual qubit. (ii) Counting the
number of generators here is trivial. It is simply $4^n-1$ as previously noticed 
from other arguments.