### 2021-11-01 documentos/proofdiagonalization.tex
Fix hard-coded [object Object] string (thanks!)

\section{Diagonalization of Choi Matrix $\mathcal{D}_N$} % {{{
\label{sec:appendix:diagonalization}
% 
\subsection{Single qubit} % {{{
In order to simplify the derivation of the relations, let us employ pairs of
binary indices instead of a single quaternary, i.e. $\alpha\to j +2 k$. For
the sake of clarity, we will use Latin symbols for binary indices, and
reserve the Greek ones for quaternary ones. 
We can write the elements of the Pauli basis
$(\sigma_{0},\sigma_{1},i\sigma_{2},\sigma_{3})$, 
compactly as
$\sigma_{k l }=\sum_{j=0}^{1} (-1)^{j k }\dyad{j}{j\oplus l }$. In vectorized form
\begin{equation}
\kket{\sigma_{k l }}=\sum_{j=0}^{1} (-1)^{j k }\kket{j,j\oplus l },
\label{CompToPau}
\end{equation}
and its inverse relation,
\begin{equation}
\kket{k,k\oplus l}=\frac{1}{2}\sum_{j=0}^{1} (-1)^{jk}\kket{\sigma_{jl}}.
\label{PauToComp}
\end{equation}
On the other hand, the matrix form of an arbitrary Pauli map $\mathcal{E}$ may
be written as
\begin{align}
\hat{\mathcal{E}} & = 
\frac{1}{2}\sum_{l m=0}^{1} \tau_{l m} \ddyada{\sigma_{l m}} 
\label{eq:inicio:canal:pauli}\\
& = \frac{1}{2}\sum_{jkl m} \tau_{l m} (-1)^{l (j+k)} 
     \ddyad{j,j\oplus m}{k,k\oplus m}.
% \label{Map_Pauli}
\end{align}
After applying the reshuffling operation on $\hat{\mathcal{E}}$, we obtain the
Choi matrix associated to the map. It reads
\begin{equation}
 \mathcal{D} = 
  \frac{1}{2}\sum_{jklm} \tau_{lm} (-1)^{l(j+k)} 
      \ddyad{j,k}{j\oplus m,k\oplus m}.
\label{Choi_a1}
\end{equation}
Note furthermore that the expression above may also be written as a combination
of tensor products of Pauli matrices
\begin{equation}
 \mathcal{D} = 
   \frac{1}{2}\sum_{l m } \tau_{l m } \sigma_{l m } \otimes \sigma_{l m }^*.
 \label{Choi_Pauli_1q}
\end{equation}
Back to the Eq. \ref{Choi_a1}), let us apply the index relabeling $k\to j\oplus
k$, then the Choi matrix reads
\begin{equation}
 \mathcal{D} =
    \frac{1}{2}\sum_{jk lm} \tau_{ lm} (-1)^{ l k} 
        \ddyad{j,j\oplus k}{j\oplus m,j\oplus m\oplus k},
\nonumber
% \label{Choi_a1}
\end{equation}
% thanks to a simplification in the argument of the power of $-1$. To continue, 
since $ (-1)^{j+j\oplus k} = (-1)^k $. To continue, we 
use the relation between computational and Pauli elements [\eref{PauToComp}],
and notice that $\sum_j(-1)^{j(m\pm n)}=2\delta_{mn}$. We arrive to the
simple expression 
\begin{align}
 \mathcal{D} & = 
  \frac{1}{2}\sum_{jk} 
      \left(\frac{1}{2}\sum_{l  m } 
(-1)^{l  k+j m }\right) 
\tau_{l  m } 
     \ddyad{\sigma_{jk}}{\sigma_{jk}}.
\label{eq:D:diagonalize}
\end{align}
Notice that  $\mathcal{D}$ is already written in its diagonal form, and one can 
identify by inspection the eigenvalues. 
% Due to the fact that we are dealing with binary indices, thus we can safely
% write the modular sums in the exponents as standard sums. Employing the
% relation between computational and Pauli elements [\eref{PauToComp}],
% noticing that $\sum_j(-1)^{j(m\pm n)}=2\delta_{mn}$, and after some steps we
% have
% \begin{align*}
%  \mathcal{D} & = 
% \frac{1}{8}\sum_{jkMnl m} \tau_{l m}(-1)^{l  M+kj+n(j\oplus m)} 
%            \ddyad{\sigma_{kM}}{\sigma_{nM}},\\
%  & = \frac{1}{4}\sum_{Mnl  m } \tau_{l  m } (-1)^{l  M+n m } 
%            \ddyad{\sigma_{nM}}{\sigma_{nM}},\\
%  & = \frac{1}{2}\sum_{Mn} 
%       \left(\frac{1}{2}\sum_{l  m } \tau_{l  m } (-1)^{l  M+n m }\right) 
%      \ddyad{\sigma_{nM}}{\sigma_{nM}},
% \end{align*}
% Then $\mathcal{D}$ is diagonal in the Pauli basis, with eigenvalues given by
The eigenvalues read
\begin{equation}
\lambda_{jk}=\frac{1}{2}\sum_{l m } 
(-1)^{jl +k m }
\tau_{l m },
% \lambda_{mn}=\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu},
% m -> j 
% n -> k
% mu -> l
% nu -> m
\label{eq:lambda:is:tau:appendix}
\end{equation}
or more compactly $\vec{\lambda} =(1/2) H \otimes H \vec{\tau}$, where 
$H$ is the Hadamard matrix.

For the sake of convenience in the demonstration of several useful properties
of the PCE channels, 
we shall reorder the eigenvalues, to write 
\begin{equation}
\lambda =  \frac12 \sa \tau
\end{equation}
with 
\begin{equation}
\sa=\left(
\begin{array}{cccc}
1 &1 &1   &1   \\
 1 & 1  & -1&-1   \\
1  & -1  & 1 & -1\\
1&-1&-1&1  
\label{a_appendix}
\end{array}
\right),
\end{equation}
instead of $H\otimes H$. This can be done due to the fact
that both matrices ($\sa$ and $H \otimes H$) are equivalent up to a permutation of rows. In other words
this operation corresponds to a relabeling in the eigenvalues $\lambda_{mn}$.
% }}}
\subsection{$N$ qubits} % {{{
To work out the $N$ qubit case, we again rely on binary indices. In this case, we replace 
$N$ dimensional vector $\valpha$ with a pair of $N$ dimensional vector binary
indices $\vec j$ and $\vec k$ so that 
each entry $\alpha_i$ of $\valpha$ is identified with the pair $j_i$ and $k_i$ as in the 
single qubit case of the previous subsection. Then, all the steps leading to 
\eref{eq:lambda:is:tau:appendix} can be redone, but with vector indices instead of binary scalars.

The tensor product of Pauli matrices, in vector form,  will be denoted by 
$\kket{\sigma_{\vec{k}\vec{l} }} = \kket{\sigma_{k_1l_1}}\otimes \cdots \otimes \kket{\sigma_{k_Nl_N}}$
and fulfill the orthogonality relation 
$\bbrakket{\sigma_{\vec{k}\vec{l} }}{\sigma_{\vec{k}'\vec{l}'}} = 
2^N \delta_{\vec{k}\vec{k}'}\delta_{\vec{l}\vec{l}'}$. 
% \begin{equation}
% \bbrakket{\sigma_{\vec{k}\vec{l} }}{\sigma_{\vec{k}'\vec{l}'}} = 2^N \delta_{\vec{k}\vec{k}'}\delta_{\vec{l}\vec{l}'}.
% \end{equation}
With this in mind, a $N$-qubit Pauli map can be written as 
% 
% 
% Analogously to the previous case, we can define an arbitrary Pauli map acting on $N$ qubits density matrices in terms of the tensor products of vectorized Pauli matrices
\begin{equation}
\hat{\mathcal{E}}_N = 
  \frac{1}{2^N}\sum_{\vec{l}\vec{m}} \tau_{\vec{l}\vec{m}} \ddyada{\sigma_{\vec{l}\vec{m} }}. 
\end{equation}
The generalization of \eref{CompToPau} and \eref{PauToComp} read
\begin{align}
\kket{\sigma_{\vec{\mu}\vec{\nu} }}  & = 
   \sum_{\vec{j}} (-1)^{\vec{j}\cdot\vec{\mu}} \kket{\vec{j},\vec{j}\oplus \vec{\nu}},\\
%
 \kket{\vec{l},\vec{l}\oplus \vec{n}} & = 
   \frac{1}{2^N}\sum_{\vec{m}} (-1)^{\vec{m}\cdot\vec{l}}\kket{\sigma_{\vec{m}\vec{n} }}.
\end{align}

% where
% \begin{widetext}
% \begin{align}
% \kket{\sigma_{\vec{\mu}\vec{\nu} }} & = \kket{\sigma_{\mu_1\nu_1}}\otimes \cdots \kket{\sigma_{\mu_N\nu_N}},\\
% & = \left(\sum_{j_1=0}^{1} (-1)^{j_1 \mu_1}\kket{j_1,j_1\oplus \nu_1}\right)\otimes\cdots \otimes \left(\sum_{j_N=0}^{1} (-1)^{j_N \mu_N}\kket{j_N,j_N\oplus \nu_N}\right),\\
% & = \sum_{\vec{j}} (-1)^{\vec{j}\cdot\vec{\mu}} \kket{j_1,j_1\oplus \nu_1}\otimes\cdots \otimes \kket{j_N,j_N\oplus \nu_N},\\
% & = \sum_{\vec{j}} (-1)^{\vec{j}\cdot\vec{\mu}} \kket{\vec{j},\vec{j}\oplus \vec{\nu}},
% \end{align}
% \end{widetext}
% and
% \begin{equation}
% \bbrakket{\sigma_{\vec{\mu}\vec{\nu} }}{\sigma_{\vec{\mu}'\vec{\nu}'}} = 2^N \delta_{\vec{\mu}\vec{\mu}'}\delta_{\vec{\nu}\vec{\nu}'}.
% \end{equation}
% 
% Recall the relation between the computational and Pauli basis elements (Eq. \ref{PauToComp}), hence for the $N$-qubits case we have
% % 
% \begin{widetext}
% \begin{align}
%  \kket{\vec{l},\vec{l}\oplus \vec{n}} & = \kket{l_1,l_1\oplus n_1} \otimes\cdots \otimes \kket{l_N,l_N\oplus n_N} \\
%   & = \left( \frac{1}{2}\sum_{m_1=0}^{1} (-1)^{m_1l_1}\kket{\sigma_{m_1n_1}}\right)\otimes\cdots \otimes\left(\frac{1}{2}\sum_{m_N=0}^{1} (-1)^{m_Nl_N}\kket{\sigma_{m_Nn_N}}\right),\\
%  & = \frac{1}{2^N}\left(\sum_{\vec{m}} (-1)^{\vec{m}\cdot\vec{l}}\kket{\sigma_{m_1n_1}}\otimes\cdots \otimes\kket{\sigma_{m_Nn_N}}\right),\\
%  & = \frac{1}{2^N}\sum_{\vec{m}} (-1)^{\vec{m}\cdot\vec{l}}\kket{\sigma_{\vec{m}\vec{n} }}.
% \label{PauToCompN}
% \end{align}
% \end{widetext}
By employing the previous relations, we can write the matrix representation of
the map, $\hat{\mathcal{E}}_N $ in the $N$-qubits computational basis as
\begin{align}
\hat{\mathcal{E}} 
% & = \frac{1}{2}\sum_{l m=0}^{1} \tau_{l m} \ddyada{\sigma_{l m}} 
& = \frac{1}{2}\sum_{\vec j\vec k\vec l \vec m} 
     \tau_{\vec l \vec m} (-1)^{\vec l (\vec j+\vec k)} 
     \ddyad{\vec j,\vec j\oplus \vec m}{\vec k,\vec k\oplus \vec m}.
% \label{eq:inicio:canal:pauli:N}
% \label{Map_Pauli}
\end{align}
% 
% 
% \begin{align}
% \hat{\mathcal{E}}_N = & \frac{1}{2^N}\sum_{\vec{\mu}\vec{\nu}} \tau_{\vec{\mu}\vec{\nu}} \ddyada{\sigma_{\vec{\mu}\vec{\nu} }},\\
% = & \frac{1}{2^N}\sum_{\vec{j}\vec{k}\vec{\mu}\vec{\nu}} \tau_{\vec{\mu}\vec{\nu}} (-1)^{\vec{\mu}\cdot(\vec{j}+\vec{k})} \ddyad{\vec{j},\vec{j}\oplus \vec{\nu}}{\vec{k},\vec{k}\oplus \vec{\nu}}.
% \end{align}
In this way it is straightforward to apply the reshuffling operation on
$\hat{\mathcal{E}}_N$ to obtain the associated Choi matrix 
then transform back to the Pauli basis  and simplify to obtain 
\begin{align}
\mathcal{D}_N & = 
   \frac{1}{2^{N}}\sum_{\vec{m}\vec{n}} \left( 
        \frac{1}{2^{N}}\sum_{\vec{l}\vec{m}} \tau_{\vec{l}\vec{m}} (-1)^{\vec{l}\cdot\vec{n}+\vec{m}\cdot\vec{m}} \right)\ddyad{\sigma_{\vec{m}\vec{n} }}{\sigma_{\vec{m}\vec{n} }}.
\end{align}
All intermediate steps, from \eref{eq:inicio:canal:pauli} to 
\eref{eq:D:diagonalize} are similar, but with a vectorized version of the 
indices, and appropriate normalization constants. 
Again, we are left with an expression that displays explicitly the eigenvalues of 
the Choi matrix, so we can write 
% \begin{equation}
% \mathcal{D}_N =\frac{1}{2^N}\sum_{\vec{j}\vec{k}\vec{\mu}\vec{\nu}} \tau_{\vec{\mu}\vec{\nu}} (-1)^{\vec{\mu}\cdot(\vec{j}+\vec{k})} \ddyad{\vec{j},\vec{k}}{\vec{j}\oplus \vec{\nu},\vec{k}\oplus \vec{\nu}}.
% \end{equation}
Thus the eigenvalues of the Choi matrix $\mathcal{D}_N$ may be written as
\begin{equation}
\lambda_{\vec{j}\vec{n}}= \frac{1}{2^{N}}\sum_{\vec{l}\vec{m}} 
    (-1)^{\vec{j}\cdot\vec{l}+\vec{k}\cdot\vec{m}}
    \tau_{\vec{l}\vec{m}} 
\end{equation}
or more compactly $\vec{\lambda} = (H\otimes H/2)^{\otimes N} \vec{\tau}$. 
Again, we prefer to reorganize the indices to be able to write
\begin{equation}
\lambda = \frac{1}{2^N} A \tau, 
\end{equation}
where 
\begin{equation}
\san=a^{\otimes N}. 
\end{equation}
% As in the case of a single qubit, the analyses in the main text employ the
% matrix $\sa$ defined in Eq. \ref{a_appendix}, to define the $\san=a^{\otimes
% N}$, instead of $\sa=H\otimes H$.

% }}}
% }}}