### 2021-11-30 documentos/images/twoQ_PCEGs_4by4_v2.tex
Another option for 2-qubit PCEGs fig

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{lscape}
\usepackage{tikz}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\newcommand{\mcG}{\mathcal{G}}

\begin{document}
\pagestyle{empty}
\begin{center}
\includegraphics[width=.5\textwidth]{twoQ_PCEGs_v2}
\begin{tikzpicture}[x=1mm,y=1mm,overlay,remember picture]
\node[inner sep=0pt] (usac) at (-88.5,77.5) {\Huge \textcolor{Red}{$\cdot$}};
\node[inner sep=0pt] (usac) at (-69.8,87.5)  {\Huge \textcolor{Red}{$\cdot$}};
\node[inner sep=0pt] (usac) at (-66,87.5) {\Huge \textcolor{Red}{$\cdot$}};
%\node[inner sep=0pt] (usac) at (-88.5,97) {\Huge .};
\node[inner sep=0pt] (usac) at (-79.15,76) {\large $\mcG_{(0,0)}$};
\node[inner sep=0pt] (usac) at (-56.65,76) {\large $\mcG_{(0,1)}$};
\node[inner sep=0pt] (usac) at (-34.15,76) {\large \textcolor{Red}{$\mcG_{(0,2)}$}};
\node[inner sep=0pt] (usac) at (-11.65,76) {\large \textcolor{Red}{$\mcG_{(0,3)}$}};
\node[inner sep=0pt] (usac) at (-75,61) {\large \textcolor{Red}{$\mcG_{(1,0)}$}};
\node[inner sep=0pt] (usac) at (-52,61) {\large \textcolor{Red}{$\mcG_{(1,1)}$}};
\node[inner sep=0pt] (usac) at (-29,61) {\large \textcolor{Red}{$\mcG_{(1,2)}$}};
\node[inner sep=0pt] (usac) at (-7,61) {\large \textcolor{Red}{$\mcG_{(1,3)}$}};
\node[inner sep=0pt] (usac) at (-75,39.5) {\large \textcolor{Red}{$\mcG_{(2,0)}$}};
\node[inner sep=0pt] (usac) at (-52,39.5) {\large \textcolor{Red}{$\mcG_{(2,1)}$}};
\node[inner sep=0pt] (usac) at (-29,39.5) {\large \textcolor{Red}{$\mcG_{(2,2)}$}};
\node[inner sep=0pt] (usac) at (-7,39.5) {\large \textcolor{Red}{$\mcG_{(2,3)}$}};
\node[inner sep=0pt] (usac) at (-75,17) {\large \textcolor{Red}{$\mcG_{(3,0)}$}};
\node[inner sep=0pt] (usac) at (-52,17) {\large \textcolor{Red}{$\mcG_{(3,1)}$}};
\node[inner sep=0pt] (usac) at (-29,17) {\large \textcolor{Red}{$\mcG_{(3,2)}$}};
\node[inner sep=0pt] (usac) at (-7,17) {\large \textcolor{Red}{$\mcG_{(3,3)}$}};
\end{tikzpicture}
\end{center}
\end{document}