### 2021-12-16 documentos/pce.tex
I was told to leave it alone, but I have this thing called OCD, you see

\section{PCE Maps} 
To introduce the family of Pauli component erasing maps we start with 
review basic concepts of quantum channels that will allow us to introduce some 
notation. 
% In this section we introduce the definition of Pauli component erasing maps,
% but first we go through several standard definitions needed to discuss them.
\subsection{Quantum channels} % {{{
\cpnote{Cambios importantes, qeu David revise}

Quantum channels are the most general linear operations that a quantum
system undergo independently of its past~\cite{zimansbook,cirac}.
The physical system under study will be associated with a Hilbert space
denoted by \hilbert, and the set of linear operators over such space will 
be denoted by $\mcB(\hilbert)$. That way, a density matrix $\rho$ of such system
 is an element of $\mcB(\hilbert)$.

% Quantum channels comprehend the most general linear operations that a quantum
% system undergo independently of its past~\cite{zimansbook,cirac}. To introduce
% their construction and properties, first let us define several objects. We will
% denote Hilbert spaces as \hilbert and $d$-dimensional Hilbert spaces as
% $\hilbert_d$. The set of linear operators over a Hilbert space \hilbert will
% be written as $\mcB(\hilbert)$, and the set of density operators as $\mcS(\hilbert)$.
%, the subset of density matrices as
%$\mcS(\hilbert)$, which is in turn a subset of the so called trace-class
%operators $\mcT(\hilbert)$, for the finite dimensional case, the one studied
%here, $\mcT(\hilbert)$ and $\mcB(\hilbert)$ coincide~\cite{zimansbook}.

The construction of quantum channels includes basically three ingredients:
linearity, trace preservation and complete positivity. Linearity is needed to
map every convex combination of density matrices into a convex combination of the
evolution of such density matrices. The
trace preserving property is required for the process $\mcE$ to be
deterministic, and reads $\tr\mcE[\rho]=\tr \rho =1$. 
Often, this condition is easy to test.
The
complete
positivity condition is needed to preserve positive semidefiniteness and handle
the non-local feature of quantum theory. 
A linear map $\mcE$ is positive if it maps density operators to density operators, \ie
if $\mcE[\rho]\geq 0$ for all density matrices $\rho$. 
% 
% $\mcE[\Delta]\geq 0 \ \ \forall \Delta \in \mcB(\hilbert)$ hermitian with $\Delta\geq 0$\cpnote{Francois, de tu nota, la $\Delta$ ya es hermitica, pues 
% decimos que es una matriz de densidad} \ddnote{Aqui hablamos de matrices
% positivas definidas en general, estuve revisando en varios libros, y parece que
% lo mas estandar es suponer que si se dice que es positiva (o negativa) definida
% o semi-definida, automaticamente se dice que es hermítica. Sin embargo hay
% definiciones mas generales que involucran matrices no hermiticas, esas eran las
% que yo tenía en mente (pero parece que no son tan conocidas). Opto por
% mencionar que $\Delta$ es hermítico.}. 
%
On the other hand, if one extends a positive map to include an ancilla, the resulting
map is not always positive. If, for all dimensions of the ancilla, such extension
results in a positive map, we say that the original map is completely
positive~\cite{}.  Quantum channels are required to be completely positive
so as to allow the proper evolution of potentially entangled states with an 
ancilla. 
% In the other hand, a map is completely positive if extending the 
%  $\id_k \otimes \mcE$ is a positive map for $k=1,2,\dots$, where $\id_k$ is the identity channel over $\mcB(\hilbert_k)$. This ensures that every positive semidefinite operator of any
% extension of the system (where the quantum channel is now a local operation) is
% mapped to a positive semidefinite operator. The trace preserving is automatically
% fulfilled for all extensions if $\mcE$ fulfills it. Therefore, any density matrix of a system
% where $\mcE$ acts locally, is sent to a valid density matrix, this is the
% substance of complete positivity.
% \flnote{No creo que lo de Bell state sea conveniente. Mejor decir algo como: E
% is totally positive if trivially extending E to the tensor product of H with an
% arbitrary dimensional ancilla space yields a positive operation.
% }\ddnote{ejecutado} 
% Concluding, quantum channels are linear completely positive trace preserving
% maps.

A simple algorithm to test the complete positivity of a quantum channel 
was developed by 
\jami{}~\cite{jamil} and Choi~\cite{choi}.
One first exploits an isomorphism that maps a channel $\mcE$ to the state
$\choi=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$, where
$\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is a 
maximally entangled state between the original system and an ancilla.
Remarkably, the map \mcE is completely positive
if and only \choi (also called dynamical matrix) is
positive-semidefinite~\cite{jamil,choi}.

%  the
% Bell state between two copies of \hilbert.

% Quantum channels can be characterized by means of the Choi-\jami{}
% isomorphism~\cite{choi,jamil}. This is, given a linear operation
% $\mcE:\mcB(\hilbert)\to \mcB(\hilbert)$, the so-called Choi matrix of it is
% $\choi=\left(\id \otimes \mcE\right)[\proj{\Omega}{\Omega}]$ (also called
% dynamical matrix), where
% $\ket{\Omega}=1/\dim(\hilbert)\sum_i^{\dim(\hilbert)}\ket{i}\ket{i}$ is the
% Bell state between two copies of \hilbert. The map \mcE is completely positive
% if and only \choi is positive-semidefinite.
%
%

% In fact,
% a linear operation is completely positive if and only if it has a Kraus
% representation~\cite{Kraus1983}.
% The minimum number of Kraus operators $K_i$
% needed to represent the channel is called \textit{Kraus rank}. 
%

% It can be seen from Kraus decomposition that quantum channels preserve
% hermiticity, this has a practical consequence when using hermitian bases, \ie{}
% let $\left\{ \Delta_i \right\}$ be a hermitian basis in $\mcB(\hilbert)$, then
% the entries of the matrix representation of the quantum channel $\mcE$, $\hat
% \mcE_{ij}=\tr\left( \Delta_i \mcE[\Delta_j] \right)$, are real. In particular,
% in this work we are interested in systems of $N$ qubits, and we use hermitian
% bases consisting of the so-called \textit{Pauli strings} \ddnote{aqui
% introduzco el nombre estandar}, defined as 
% \begin{equation}
% \sigma_{\vec \alpha}:=\sigma_{\alpha_1}\otimes \sigma_{\alpha_2} \otimes \dots \otimes \sigma_{\alpha_N},
% \label{eq:pauli_strings}
% \end{equation}
% where $N$ is the number qubits, $\alpha_i=0,1,2,3$ with $\sigma_0:=\one$ and
% $\sigma_{1,2,3}:=\sigma_{x,y,z}$. Thus, given $N$ and the fact that $\tr
% \sigma_{\vec \alpha}=\delta_{\vec\alpha,\vec 0} 2^{N}$, it is clear that the
% basis $\left\{ 2^{N/2} \sigma_{\vec \alpha} \right\}$ is orthonormal in
% $\mcB(\hilbert_{2^N})$. The case with $N=1$ corresponds to the well-known
% \textit{Pauli basis},
% $1/\sqrt{2} \left\{ \one, \sigma_x,\sigma_y,\sigma_z \right\}$.

%

% \flnote{
% probablemente convenga mostrar la ecuaci'on desplegada. Asimismo se debe decir que 
% $\sum_i K_i^\dagger K_i = 1$ para la trace-preserving property. 
% Creo qie ``hermiticity preserving'' es en parte consecuencia de la definición, pero esto no es muy importante
% }\ddnote{ejecutado, hacia falta edicion, sin embargo lo de hermiticity preserving es para hacer incapie en por que podemos trabajar en una representacion totalmente real de los canales.}
% 
%   \ddnote{[Choi representation, Kraus operators and Stinespring dilation]}

% }}}
\subsection{Pauli component erasing maps} % {{{
\label{sec:single:qubit}
% \todo{David escribe una primera version y luego Carlos itera}

We shall start exploring a single qubit scenario, where the most general 
density matrix can be written as
\begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
\rho=\frac12 {\sum_{\alpha=0}^3 r_\alpha \sigma_\alpha},
\end{equation}
with $\sigma_0= \one$, and $\sigma_{1,2,3}$ the usual Pauli matrices. 
Normalization requires that $r_0=1$ and the remaining $r_{1,2,3}$ 
form the Bloch vector. Consider the map that projects
each component 
in the following way,
\begin{equation}
r_\alpha \mapsto \tau_\alpha r_\alpha
\end{equation}
where $\tau_\alpha$ is either $1$ or $0$. Not every
operation like this is a quantum map; for example collapsing the entire Bloch
ball to the $xy$ plane (setting $\tau_3=0$ and the rest to 1) is not a quantum
channel. A direct evaluation of the complete positivity
conditions~\cite{zimansbook,davalosdivisibility}, 
\begin{align}
1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
1+\tau_\alpha-\tau_\beta-\tau_\gamma &\geq 0 \ \ \forall \alpha\neq \beta \neq \gamma,
\end{align}
where trace-preserving is already imposed ($\tau_0=1$), show that five out of
the eight 1-qubit PCE maps are quantum channels.
% 
% show that only five
% operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
These operations are the identity map, the completely depolarizing channel
($\rho \mapsto \one/2$); and completely bit, phase, and the bit-phase flip
channels~\cite{chuangbook}, 
% These operations are by definition diagonal in the
% Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat
% \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$
and can be pictured using one column tables showing the positions of $0$s and
$1$s, see~\fref{fig:one:qubit:examples}.

% \pce{} maps are defined as those quantum channels\janote{David, los mapas PCE
% no son necesariamente canales cuánticos. Por esto, estás siendo incosistente en
% el uso de ``\pce{} maps'' en las líneas \lr{A1}, \lr{A2} y \lr{A3} del PDF.
% Especialmente, en la línea \lr{A1} seguro te estás refiriendo a todos los
% mapas, no sólo a los canales cuánticos} that \textit{nullify} Pauli strings.
% \janote{Por el comentario que hice antes propongo la siguiente reformulación:} 
% \janote{
% \pce{} maps are defined as positive trace-preserving (PTP) maps that 
% \textit{nullify} Pauli strings.}
% In the case of one qubit, they consist of maps nullifying the averages
% $r_\alpha:=\langle \sigma_\alpha \rangle=\tr \left( \rho \sigma_\alpha \right)$
% with $\alpha=1,2,3$. The three components $r_\alpha$ form the well-known
% \textit{Bloch vector} $\vec r$, describing one qubit density matrices,
% \begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
% \end{equation}
% Therefore, maps erasing components of $\vec r$ are parametrized in the following way,
% \begin{equation}
% r_\alpha \mapsto \tau_\alpha r_\alpha
% \end{equation}
% where $\tau_\alpha$ is either $1$ or $0$. It is well-known that not every operation like this is a quantum map, for example collapsing the entire Bloch ball into a disk of radio equal to one is not a quantum channel. A direct evaluation of the complete positivity conditions~\cite{zimansbook,davalosdivisibility},
% \begin{align}
% 1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
% 1+\tau_i-\tau_j-\tau_k &\geq 0 \ \ \forall i\neq j \neq k,
% \end{align}
% where trace-preserving is already imposed ($\tau_0=1$), show that only five operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
% the identity map, the completely depolarizing channel ($\rho \mapsto \one/2$); and completely bit, phase, and the bit-phase flip channels~\cite{chuangbook}. These operations are by definition diagonal in the Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$, and can be pictured using one column tables showing the positions of $1$s and $0$s in the diagonal of the channel (in the Pauli basis), see~\fref{fig:one:qubit:examples}.
\begin{figure} % {{{
\begin{center}
\begin{tikzpicture}[x=0.5cm, y=0.5cm] % {{{
\pgfmathsetmacro{\unitstep}{3.5}
% Coordenadas   {{{
\node at (-0.6,0.5) {(a)} ;
\draw[black!10] (0,0) rectangle (1,1); \node at (0.5,0.5) {$\tau_0$} ;
\begin{scope}[shift={(0,-1)}]
\draw[black!10] (0,0) rectangle (1,1); \node at (0.5,0.5) {$\tau_1$} ;
\end{scope}
\begin{scope}[shift={(0,-2)}]
\draw[black!10] (0,0) rectangle (1,1); \node at (0.5,0.5) {$\tau_2$} ;
\end{scope}
\begin{scope}[shift={(0,-3)}]
\draw[black!10] (0,0) rectangle (1,1); \node at (0.5,0.5) {$\tau_3$} ;
\end{scope} 
\draw (0,-3) rectangle (1,1); 
% }}}
\begin{scope}[shift={(1*\unitstep,0)}] % Identity {{{
% \begin{scope}[shift={(\unitstep*2,0)}]
\node at (-0.6,0.5) {(b)} ;
\fill[black] (0,0) rectangle (1,1);
\draw[black!10] (0,0) rectangle (1,1);
\begin{scope}[shift={(0,-1)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\draw (0,-3) rectangle (1,1); 
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,0)}] % Dephasing {{{
\node at (-0.6,0.5) {(c)} ;
\fill[black] (0,0) rectangle (1,1);
% \draw (0,0) rectangle (1,1);
\draw[black!10] (0,0) rectangle (1,1);
\begin{scope}[shift={(0,-1)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\draw (0,-3) rectangle (1,1); 
% \begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \draw (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \draw (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \draw (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(3*\unitstep,0)}] % Depolarization {{{
\node at (-0.6,0.5) {(d)} ;
\fill[black] (0,0) rectangle (1,1);
% \draw (0,0) rectangle (1,1);
\draw[black!10] (0,0) rectangle (1,1);
\begin{scope}[shift={(0,-1)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\draw (0,-3) rectangle (1,1); 
% \begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \draw (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \draw (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \draw (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(4*\unitstep,0)}] % (e) canal malo {{{
\node at (-0.6,0.5) {(e)} ;
\fill[black] (0,0) rectangle (1,1);
% \draw (0,0) rectangle (1,1);
\draw[black!10] (0,0) rectangle (1,1);
\begin{scope}[shift={(0,-1)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \draw[black!10] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \draw (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \draw (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \draw (0,0) rectangle (1,1); \end{scope}
\draw (0,-3) rectangle (1,1); 
\end{scope} % }}}
\end{tikzpicture} % }}}
\end{center}
	\caption{In (a) we introduce the positions of the single 
qubit diagrams, so that each square corresponds to a single 
$\tau_{\alpha}$, $\alpha=0,1,2,3$. The diagrams in 
(b), (c) and (d) correspond to the identity map, dephasing channel, and 
total depolarization, respectively. In (e) we show a map that only erases the component
$r_1$, collapsing the Block sphere into a disk, and thus does not correspond to a quantum operation.}
	\label{fig:one:qubit:examples}
\end{figure} % }}}
% }}}
% \subsection{$\mathbf{N}$ qubits Pauli component erasing maps} % {{{

In order to present and work the $N$ qubit case, it is useful 
to introduce 
the so-called \textit{Pauli strings}, defined as 
\begin{equation}
\sigma_{\vec \alpha}=\sigma_{\alpha_1}\otimes \sigma_{\alpha_2} \otimes \dots \otimes \sigma_{\alpha_N},
\label{eq:pauli_strings}
\end{equation}
where $\vec \alpha$ denotes a vector index $\qty(\alpha_1,\ldots,\alpha_N)$
and $\alpha_i=0,1,2,3$. These hermitian operators form an orthogonal 
basis in the space of operators acting on $N$ qubits. In fact, 
$\tr \sigma_{\vec \alpha} \sigma_{\vec \alpha'}=2^N\delta_{\alpha \alpha'}$
and 
$\tr \sigma_{\vec \alpha}=\delta_{\vec\alpha,\vec 0} 2^{N}$.
% 
% Thus, given $N$ and the fact that $\tr
% \sigma_{\vec \alpha}=\delta_{\vec\alpha,\vec 0} 2^{N}$, it is clear that the
% basis $\left\{ 2^{N/2} \sigma_{\vec \alpha} \right\}$ is orthonormal in
% $\mcB(\hilbert_{2^N})$. The case with $N=1$ corresponds to the well-known
% \textit{Pauli basis},
% $1/\sqrt{2} \left\{ \one, \sigma_x,\sigma_y,\sigma_z \right\}$.

Similarly to the one qubit case, the density matrix $\rho$ of a system of $N$
qubits can be written using Pauli strings in the following way, 
\begin{equation}\label{eq:N_qubits_rho}
	\rho =\frac{1}{2^N} 
            \sum_\valpha r_\valpha \sigma_\valpha,
%             \quad r_{\vec0}=1,
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
%             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
	%r_{0,\ldots,0}=1,
\end{equation}
so $r_{\vec \alpha}=\langle \sigma_{\vec \alpha} \rangle=\tr \left(\rho
\sigma_{\vec \alpha} \right)$ is the coefficient corresponding to the expansion
of the density matrix in the normalized basis of Pauli strings. Again,
normalization of the state requires that $r_{\vec 0}=1$.
We shall refer to $\paulicomponents$ as the ``Pauli components'' of the density
matrix of a system of qubits.

% The components of $\valpha$ range from $0$ to $3$.  Notice that
% given that $\tr \rho=1$ and $\sigma_{\vec 0} = \openone$, it is clear that
% $r_{\vec 0}=1$. Additionally, observe that $\vec r_{\vec \alpha}$ can be
% understood as a generalization of the Bloch vector for $N$ qubits.
%

% The density matrix $\rho$ of a system of $N$ qubits can be written in Pauli
% matrices basis as 
% \begin{equation}\label{eq:N_qubits_rho}
% 	\rho =\frac{1}{2^N} 
%             \sum_\valpha r_\valpha \sigma_\valpha,
% %             \quad r_{\vec0}=1,
% % 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
% %             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% % 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
% 	%r_{0,\ldots,0}=1,
% \end{equation}
% where $\vec \alpha$ denotes a vector index $\qty(\alpha_1,\ldots,\alpha_N)$,
% $\sigma_{\vec \alpha} := \sigma_{\alpha_1}\otimes \ldots \otimes
% \sigma_{\alpha_N}$ is the tensor product of Pauli matrices, and 
% $r_{\vec \alpha}$ is the coefficient corresponding to the expansion
% of the density matrix in the complete orthonormal basis $\left\{ 2^{-N/2}\sigma_{\vec
% \alpha} \right\}$. In the sum each of the $N$ components of $\valpha$, i.e.
% $\alpha_{1,\cdots,N}$ is varied from $0$ to $3$. 
% Notice that $\sigma_{\vec 0} = \openone$, and since both $\tr \rho=1$ and 
% $\tr \sigma_\valpha=2^N\delta_{\valpha,\vec 0}$, then $r_{\vec 0}=1$. 
% We shall refer to $\paulicomponents$ as the ``Pauli components'' of the density
% matrix of a system of qubits.

A Pauli component erasing (PCE) operation is a Pauli diagonal operation
that either preserves or completely erases the Pauli components of a
density matrix. That is, 
% it transforms its Pauli components as 
\begin{equation} \label{eq:PCE_definition}
% \begin{split}
% 	\paulicomponents \mapsto \ \taus \paulicomponents,\hspace{.5cm}\\
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
	\paulicomponents \mapsto \taus \paulicomponents,\quad 
	\taus = 0,1.
% \end{split}
\end{equation}
It is worth noticing that not all PCE maps are quantum operations. For single 
qubits, if just one of the $\taus=0$, the complete positivity condition 
is not met, as discussed in \sref{sec:single:qubit}. It will be of central interest
in this article to determine which $\taus$ lead to a valid quantum map.

As mentioned above, \pce{} maps\R{A1} nullify Pauli strings, thus, consider the linear operations acting in the generalized Bloch vector $\vec r_{\vec \alpha}$,
\begin{equation}
	\paulicomponents \mapsto \taus \paulicomponents.
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1,
\label{eq:definition:Pauli}
\end{equation}
with $\tau_{\vec \alpha}$ either $1$ or $0$.
\cpnote{Revisar aca lo que sigue}
Similar to the qubit case, not every operation with this form is a completely
positive trace perserving map. For the operation to be trace preserving, it is
easy to prove that $\tau_{\vec 0}=1$. On the other hand, constructing and
evaluating the conditions for complete positivity is involved and is the main
problem addressed in this paper. Moreover, these operations are by definition
diagonal in the $\left\{ 2^{-N/2}\sigma_{\vec
\alpha} \right\}$ basis.

\cpnote{Creo qeu vale la pena mencionar la relacion con los canales de Ruskai. 
Lo discutimos en pleno, o por lo menos cuando esten JA y FL. quizá tambien discutir
si se mencionan a los canales  \cite{Siudzinska2020}}\ddnote{Eh estado trabajando en esto, pero el material que he encontrado va mejor en la introduccion, y en todo caso, creo que es muy pronto para mencionar detalles tecnicos sobre los canales que coinciden con los de Ruskai, pues ni siquiera hemos dicho que operaciones de las que definimos aqui, son válidas}

It is worth noticing that the set of \pce{} maps \R{A2} overlaps with the set of ``Pauli diagonal channels constant on axes'' defined in Ref.~\cite{nathanson2007pauli}, consisting of convex combinations of \textit{quantum-classical} channels. In particular, it can be shown that quantum-classical channels defined with the eigenbasis of some set of $2^N-1$ commuting Pauli observables~\cite{Lawrence2002}, is a \pce{} map with exactly $ 2^N $ components equal to $1$s in its diagonal.

\cpnote{davalos, dale aca un pequeño rollo relevante quiza de una o dos
frases de la importancia de esta familia.} \ddnote{hecho, pero de pronto quedo espeso. Me estan gustando las cosas que he estado encontrando, voy a investigar en esa direccion por mi cuenta, si sale algo interesante, los invito, como no}

%\ddnote{hay un typo en esta ecuacion}
% A PCE operation is a trace-preserving operation that leaves invariant or erases 
% the Pauli components of a density matrix of $n$ qubits. \janote{Ojo que aquí 
% habrá que unificar la notación de subíndices. No lo hago porque no lo tengo del 
% todo claro y no hay problema en dejarlo así por el momento.}
% \cpnote{Aca voy}

For the two qubits case, the operations illustrated
in~\eref{eq:definition:Pauli} can be pictured using $4\times 4$ tables, showing
the positions of the $0$s and $1$s of the diagonal using the Pauli string
basis, see~\fref{fig:two:qubit:examples}. This way of representing \pce{}
maps\R{A3} turned out to be very useful for their analysis throughout this
work.

A graphical representation for PCE maps may be introduced, being the
two qubits case what has proved to be the most useful. Consider a $N$ dimensional 
Cartesian grid, with $4^N$ places.  Each place has $N$ integer coordinates, ranging
from 0 to 3, so each place corresponds to a given $\vec \alpha$ in 
\eref{eq:N_qubits_rho}. For a given PCE, we shall fill 
the square if the corresponding $\tau_{\vec \alpha}=1$. Otherwise, we leave it 
empty. Examples for $N=1$ are provided in \fref{fig:one:qubit:examples}, whereas
for two qubits, we exemplify them in \fref{fig:two:qubit:examples}.

% Notice
% than in both cases we drop the label $\alpha$ for simplicity, as it will 
% remain unchanged throughout this text \janote{No entiendo qué estás diciendo
% aquí}.

\begin{figure} % {{{
\begin{tikzpicture}[x=0.55cm, y=0.55cm] % {{{
\pgfmathsetmacro{\unitstep}{5.2}
\begin{scope}[shift={(0*\unitstep,0)}] % Coordenadas {{{
\node at (-0.5,0.5) {(a)} ;
% \cuadritotikz{}
    \foreach \x in {0,1,2,3} {
      \foreach \y in {0,1,2,3} {
        \begin{scope}[shift={(\x,-\y)}] 
          \draw[black!10] (0,0) rectangle (1,1); 
%           \draw (0,0) rectangle (1,1); 
          \node at (0.5,0.5) {\scalebox{.8}{$\tau_{(\y,\x)}$}};
         \end{scope}
%         \node at (0,-\y) (input\y) {$i_\y$};
%         \node[block] at (2,-\y) (block\y) {$f_\y$};
%         \draw[->] (input\y) -- (block\y);
%         \draw[->] (block\y.east) -- +(0.5,0);
    }
    }
 \draw (0,-3) rectangle (4,1);
\end{scope} % }}}
\begin{scope}[shift={(1*\unitstep,0)}] % Good channel {{{
\node at (-0.5,0.5) {(b)} ;
\cuadritotikz{}
%     \foreach \x in {0,1,2,3} {
%       \foreach \y in {0,1,2,3} {
%         \begin{scope}[shift={(\x,-\y)}] 
%           \draw (0,0) rectangle (1,1); 
% %           \node at (0.5,0.5) {$\tau_{\y,\x}$};
%          \end{scope}
% %         \node at (0,-\y) (input\y) {$i_\y$};
% %         \node[block] at (2,-\y) (block\y) {$f_\y$};
% %         \draw[->] (input\y) -- (block\y);
% %         \draw[->] (block\y.east) -- +(0.5,0);
%     }
%     }
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,0)}] % Good channel {{{
\node at (-0.5,0.5) {(c)} ;
\cuadritotikz{}
%     \foreach \x in {0,1,2,3} {
%       \foreach \y in {0,1,2,3} {
%         \begin{scope}[shift={(\x,-\y)}] 
%           \draw (0,0) rectangle (1,1); 
% %           \node at (0.5,0.5) {$\tau_{\y,\x}$};
%          \end{scope}
% %         \node at (0,-\y) (input\y) {$i_\y$};
% %         \node[block] at (2,-\y) (block\y) {$f_\y$};
% %         \draw[->] (input\y) -- (block\y);
% %         \draw[->] (block\y.east) -- +(0.5,0);
%     }
%     }
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\end{tikzpicture} % }}}
	\caption{In (a) we introduce the
positions of two qubit diagrams. The diagrams in 
(b) corresponds to a quantum channel that results from the tensor product of
bit flip channels in each qubit [see \fref{fig:one:qubit:examples}(c)],
and in (c) a diagram of a map that is not a quantum channel is presented.}
	\label{fig:two:qubit:examples}
\end{figure} % }}}

% In \Fref{fig:PCE_figs} we introduce a pictorial representation of PCE operations of 
% 1, 2 and 3 qubits. Every position in the column, board or cube is associated with 
% the subindices of $\taus$ of a PCE operation. If a little square or cube  in position
% $\alpha_1,\ldots,\alpha_N$ is black, then $\taus=1$, if it is white, then $\taus=0$. For example,
% \begin{center}
% \includegraphics[width=2cm]{2qubits_pce_example}
% \end{center}
% this represents a PCE operation of 2 qubits that leaves invariant the Pauli 
% components $r_{0,0}$, $r_{0,2}$, $r_{1,0}$, $r_{2,2}$, $r_{3,2}$. This operation
% is not completely positive and therefore is not a quantum channel. In contrast, 
% \begin{center}
% \includegraphics[width=2cm]{2qubits_pceChannel_example01}
% \end{center}
% this other example of PCE operation that leaves invariant Pauli components 
% $r_{0,0}$, $r_{1,2}$, $r_{2,1}$, $r_{3,3}$ satisfies complete positivity and 
% is a quantum channel. 
% 
%

% }}}