### 2022-01-19 documentos/vector_space.tex
stuff

\subsection{PCE channels as vector spaces} % {{{
\cpnote{Ojo, cambie el nombre de la seccion. Pendiente de sus comentarios}
% \ddnote{propongo channels as vector spaces como titulo de seccion, pues lo
% actual se entiende mas en el sentido estandar: canales como miembros de un
% espacio vectorial, que tambien es cierto, pero no es el punto aqui}
% \cpnote{Falta lo de las clases de equivalencias, pero aun no encuentro el
% espacio. Quizá cuando haga falta se puede poner}
% \janote{Se me hace que podemos agregar algo al final, ya que es la parte en la que
% se hablan de la regla $2^k$ y regla espejo. No tengo una idea clara, pero si te 
% parece una buena idea podría elaborar más cómo meterlo ahí}
% \par
% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{

In this subsection we will provide a one-to-one relation between PCE quantum
channels and the subspaces of a discrete vector subspace associated with the
indices $\valpha$ labeling the components of a state; see
\eref{eq:N_qubits_rho}. \cpnote{Some established facts about vector spaces will allow
us to draw conclusions for PCE quantum channels.}

Let us start by recalling that the problem of determining complete positivity
of a PCE can be recast as determining which vectors $\vtau$ are mapped via
$\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to a vector with positive entries $\lambda_\valpha$, as in \eref{eq:lambda:is:A:tau}. 
% . In components, this would read 
% \begin{equation}
% % \left(
% \sum_\vbeta
% \san_{\valpha\vbeta} \,
% % \right)
% \tau_{\vec \beta}=2^Np_{\vec\alpha}
% \label{eq:A:tau:is:p}
% \end{equation}
% where
% $\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
% \sa_{\alpha_N\beta_N}$. 
Using the fact that $a^{-1}=a/4$, so 
% (so $A = A^{-1}/2^N$) 
\begin{equation}
A^{-1} = \frac{1}{4^N} A,
\end{equation}
we can directly invert \eref{eq:lambda:is:A:tau} to obtain 
\begin{equation}
% \tau_{\vec \alpha}= \san_{\vec\alpha\vec\beta} \, p_\vbeta,
% \tau = 
% \san \lambda =2^N \tau
\sum_\vbeta \san_{\valpha \vbeta} \lambda_\vbeta =2^N \tau_\valpha
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. 
This is a remarkable equation, as it provides a method to diagonalize the
Choi matrix of any 
Pauli diagonal operation.

Other two simple but crucial
observation are the following. For valid quantum channels, 
% \begin{equation}
% % \sum_{
% \end{equation}
% $\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
% $\lambda_\valpha=0$ for all $\valpha$ in an arbitrary subset of 
% multi-indices $\Omega$, 
\begin{equation}
\sum_{\valpha \in \Omega} \lambda_\valpha=0 
\quad \implies 
\quad 
\lambda_\valpha=0, \, \forall \valpha \in \Omega
\label{eq:lambda:is:zero}
\end{equation}
for an arbitrary subset of multi-indices $\Omega$, 
as each member of the sum is
greater or equal to zero, 
due to complete positivity of the underlying channel. 
Finally, setting $\valpha =0$ in \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha)=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
% If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
If we now assume that $\tau_{\vec\alpha}=1$, and calculate the difference between
\eref{eq:sum:alpha:is:2N} and
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$
(which follows from \eref{eq:tau:is:A:p} and $\tau_{\vec\alpha}=1$), one obtains
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three
multi-indices, $\valpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$.

Let us now work out the nature of the aforementioned connection.  Let the
components of  $\vbeta$ be such that $\beta_l = \beta \delta_{lk}$ for a
particular particle index $k$, with $\beta$ such that $\sa_{\alpha_k\beta}=1$.
Since $\sa_{\alpha 0}=1$ for any $\alpha$, this particular choice of $\vbeta$
belongs to  $\Phi(\vec\alpha)$, so that \eref{eq:coneccion:dos:indices:A}
indeed holds. However, for such election of $\vbeta$ the above relation reduces
to 
\begin{equation}
\sa_{\beta\gamma_k}=\sa_{\beta\gamma_k^\prime}
\label{eq:14}
\end{equation}
for all $\beta$ such that $\sa_{{\alpha_k}\beta}=1$.  One can verify, by
working out the different cases, that \eref{eq:14} is equivalently expressed as 
\begin{equation}
\gamma_k^\prime=\alpha_k\oplus{}\gamma_k
\label{eq:oplus}
\end{equation}
where the $\oplus{}$ denotes the operation of the Klein group; see Figure
\ref{tab:1} for a detailed description. 
\par
% 
% 
% % Since all elements of $A$ are $\pm1$, and using that $A_{\valpha \vbeta}=1$
% % of $\vbeta \in \Phi(\vec\alpha)$, we may rewrite \eref{eq:coneccion:dos:indices:A}
% % as
% % \begin{equation}
% % \san_{\vbeta\vec\gamma} \san_{\vbeta\vec{\gamma^\prime}} 
% % = 
% % 1
% % , \quad
% % \forall \vec\beta\in\Phi(\vec\alpha),
% % \label{eq:12}
% % \end{equation}
% % which can be further decomposed in 
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}
% %   \ldots
% % %   \cdot\ldots\cdot 
% % \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% % \label{eq:13}
% % \end{equation}
% % Moreover, since 
% % 
% % 
% % 
% % Since $\sa_{\alpha\beta}=\pm1$, we may rewrite (\ref{eq:11}) as
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% % \label{eq:12}
% % \end{equation}
% % This must hold for all $\vec\beta\in\Phi(\vec\alpha)$. From this follows that
% % (\ref{eq:12}) may be rewritten as
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
% % \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}.
% % \label{eq:13}
% % \end{equation}
% % Here we show that (\ref{eq:13}) implies (\ref{eq:14}). 
% Since \eref{eq:13} holds for 
% all $\vec\beta\in\Phi(\vec\alpha)$, we may choose a specific such $\vec\beta$, namely
% $\beta_l=0$ for all $l\neq k$ and $\beta_k=\beta$, where $\beta$ is an arbitrary number
% such that
% $\sa_{\alpha_k\beta}=1$. 
% % \begin{equation}
% % \sa_{\alpha_k\beta}=1.
% % \end{equation}
% This $\vec\beta$, as is readily verified, belongs to $\Phi(\vec\alpha)$, so
% that \eref{eq:13} indeed holds for this $\vec\beta$. On the other hand, it is
% then readily verified that, for this $\vec\beta$, \eref{eq:13} reduces to 
% % (\ref{eq:14}), whcih therefore follows. 
% % \flnote{Leve modificaci\'on de las ecuaciones anteriores, y adecuaci\'on del
% % texto siguiente. } \cpnote{francois, recuerdo 
% % que esto lo discutimos. No se si en el momento me convencí, pero ahora no veo 
% % proque se tiene que cumplir para los indices individulaes}. 
% % We now show in Appendix
% % \ref{app:proof} that from this follows that
% % for all $1\leq k\leq N$, $\gamma_k$
% % and $\gamma_k^\prime$ are connected by
% \begin{equation}
% \sa_{\beta\gamma_k}\sa_{\beta\gamma_k^\prime}=1
% \label{eq:14}
% \end{equation}
% for all $\beta$ such that $\sa_{{\alpha_k}\beta}=1$. 
% One can verify, by working out the different cases, 
% that \eref{eq:14} is equivalently expressed as 
% \begin{equation}
% \gamma_k^\prime=\alpha_k\oplus{}\gamma_k
% \label{eq:oplus}
% \end{equation}
% where the $\oplus{}$ denotes the operation of the Klein group; see Figure
% \ref{tab:1} for a detailed description. 
% \par
% }}}
% Espacio vectorial {{{

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$.
We notice that the Klein group described in \fref{tab:1} is actually isomorphic
to the two dimensional vector space over the field of two elements $\{0,1\}$.
Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, and say that for valid quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preserved and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

\begin{figure} % {{{
\begin{center}
\begin{tabular}{| c | c c c c |}
\hline
$\oplus{}$ & 0 & 1& 2 & 3\\
\hline
0 & 0 & 1 & 2 & 3\\
1 & 1 & 0 & 3 & 2\\
2 & 2 & 3 & 0 & 1 \\
3 &3 & 2 & 1 & 0\\
\hline
\end{tabular}
\caption{Table for the $\oplus{}$ operation defined in (\ref{eq:14}). Note that
the operation is an abelian group, in fact it corresponds to the 
Klein group, where the neutral element is $0$. This is the reason for choosing
an additive notation for the operation defined in (\ref{eq:14}). }
\label{tab:1}
\end{center}
\end{figure} % }}}

% % Lo que quiero quitar {{{
% \cpnote{Propongo quitar desde aca}
% Note further that if we perform the identification
% \begin{subequations}
% \begin{eqnarray}
% 0&\to&(0,0)\label{eq:identa}\\
% 1&\to&(0,1)\label{eq:identb}\\
% 2&\to&(1,0)\label{eq:identc}\\
% 3&\to&(1,1)\label{eq:identd}
% \end{eqnarray}
% \label{eq:ident}
% \end{subequations}
% then the $\oplus$ operation reduces to simple vector addition in
% binary arithmetic. If we therefore denote by $\overline{\alpha}$
% \cpnote{I dont think this is necesary. specially if we will write something
% else}
% the multi-index $\vec\alpha$ converted to a binary vector by the componentwise
% application of (\ref{eq:ident}),
% the connection between $\vec\alpha$, $\vec\gamma$, and $\vec{\gamma^\prime}$ defined by (\ref{eq:11}) is rewritten 
% as
% \begin{equation}
% \overline\gamma=\overline\alpha+\overline{\gamma^\prime}
% \label{eq:17}
% \end{equation}
% where the addition sign here refers to ordinary binary vector addition. 
% \cpnote{Hasta aca}
% % }}}

From this readily follows an amusing property: the set of all multi-indices
$\vec\gamma$ for which $\tau_{\vec\gamma}=1$, is closed under binary
vector addition, in other words, it forms a vector subspace of the set
of all multi-indices.
A moment's consideration will further show that the above reasoning can be
inverted, that is, that if we set all $\tau_{\vec\gamma}$ equal to one
whenever the $\vec\gamma$ belong to a given vector subspace of the set of
all indices, then the vector $\vec\tau$ indeed has an image
which is a vector
with only positive components. 
In other words, there is a one to one correspondence between 
a valid quantum channel, and a vector subspace of the aforementioned space. 
% \cpnote{Aca quizá que Jose Alfredo ponga un ejemplo de una 
% figura en la que se tenga algo que es un canal y algo que no es  un canal. 
% Quizá también se puede mencionar qeu el subespacio mas chico es el canal totalmente 
% depolarizante y el subespacio mas grande es la identidad} 
\par

% }}}
% Generacion de soluciones {{{
With this information, we present a procedure to generate all solutions: we start out
from the solution having $\tau_{\vec0}=1$, with everything else $0$. We may
then successively switch $\tau$'s to $1$ for various values of $\valpha$,
taking care immediately to set equal to one the components of $\tau$ that
correspond to values of $\vbeta$ generated by the previously switched values of
$\valpha$ via the operation $\oplus$. Doing so, in an ordered way, allows one
to generate
all valid quantum channels with a given set of preserved components, without the need
of exploring the exponentially large space of PCE.  \par 
% }}}}

\begin{figure} % {{{
\begin{center}
% Del dcumento francois.pdf pg 8 agarramos:
% e11, e4 y E7
\begin{tikzpicture}[x=0.55cm, y=0.55cm] % {{{
\pgfmathsetmacro{\unitstep}{5.2}
\begin{scope}[shift={(1*\unitstep,0)}] % Depolirizng channel {{{
\node at (-0.5,0.5) {(a)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,0)}] % Interseccion {{{
\node at (-0.5,0.5) {(b)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(3*\unitstep,0)}] % E4 {{{
\node at (-0.5,0.5) {(c)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(1*\unitstep,-\unitstep)}] % E7 {{{
\node at (-0.5,0.5) {(d)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,-\unitstep)}] % E11 {{{
\node at (-0.5,0.5) {(e)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(3*\unitstep,-\unitstep)}] % Identity channel {{{
\node at (-0.5,0.5) {(f)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\end{tikzpicture} % }}}
\end{center}
\caption{
\cpnote{Revisar, solo tiene una iteracion por parte mia}
Examples of diagrams for several PCE quantum channels. In
 (a) the totally depolarizing channel is represented whereas
in (b) a PCE channel that preserves four components.  (c), (d)
and (e) show three generators; the combination of any two of them
yields the channel in (b). (f) represents the identity map. 
% 
% \cpnote{Mejorar el caption cuando me ponga de acuerdo con la gente}
% Aca poner 6 PCEs. (f) Identidad, (a) totally depolarizing, (c, d y e) tres
% generadores, (b) un canal que resulte del generador. Sobra un espacio para
% ilustrar alguna otra cosa. Quiza poner otro generador para que se vea que no
% hay una manera unica de generar esos bichos.
}
\label{fig:examplesPCE}
\end{figure} % }}}
% Los canales tiene 2^K componente {{{

% **********************************
% 
% We can show that all PCE quantum channels preserve $2^K$ components. 
% In
% fact, with the above picture, it is clearly revealed that the vector space has
% dimension $2N$. Moreover, 
% consider any subspace and chose the maximum number of linearly independent
% elements (say $K$).  Any of the  $2^K$ subsets of those elements will yield a
% unique element. Moreover, since $\valpha + \valpha = \vec 0$, these are all
% elements of the vector space. This means that any PCE quantum channel will preserve $2^K$ 
% components, for $K$ an integer. See \fref{fig:examplesPCE} for some examples. 
% \par 
% 
% **********************************
% **********************************

We can show that all PCE quantum channels preserve $2^K$ components. 
First recall that a vector space of dimension $d$ over a field of $q$ elements
has $q^d$ elements~\cite{Roman2008}.
Now $V$ is a vector space on a field of two elements having dimension $2N$. 
% The dimension of the vector space $V$ of all multi-indices is $2N$, and 
We have
seen earlier that
\begin{equation}
W=\left\{
\vec\alpha:\vec\alpha\in V, \tau_{\vec\alpha}=1
\right\}
\label{eq:19}
\end{equation}
is a subspace of $V$. 
As such, $W$ has a given dimension $K$, which means that
$W$ has $2^K$ elements. In other words, a vector $\vec\tau$ with the property
discussed above can only have $2^K$ elements equal to 1, for a given integer
$K$. \par

% **********************************
% **********************************
% 
% \cpnote{Dejar que la raza elija uno de los dos parrafos anteriores}\janote{El
% primero (simón, cambié de opinión)} \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCEs exist that conserve $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}\label{eq:conteo:simmetry}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest a relation  between individual channels that
preserve $K$ and $2N-K$ Pauli components that for the time being
has escaped our efforts to identify. 
\par % }}}

% }}}