### 2022-01-20 documentos/quantum_classical.tex
agrego archivos del apendice

\section{Quantum-Classical channels}
\label{quantum_classical}
A Quantum-classical channels is defined using orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ an orthonormal basis in $\hilbert_{2^N}$