### 2022-01-25 documentos/quantum_classical.tex
secc V

\section{Quantum-Classical channels}
\label{quantum_classical}
A Quantum-classical (CQ) channel is defined  is defined using orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ a orthonormal basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i}.
\end{equation}