### 2022-01-26 documentos/quantum_classical.tex
etc

\section{Quantum-Classical channels}
\label{quantum_classical}
\ddnote{Alguien revise este apendice, tiene solo una iterada}
A Quantum-classical (QC) channel is defined using an orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ such a basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i},
\end{equation}
this is, QC channels project density matrices onto the corresponding diagonal matrix in the basis $B$~\cite{nathanson2007pauli}.

Consider now that the basis $B$ is the simultaneous eigenbasis of a maximal set of commuting Pauli strings denoted by $\text{set}(B)$, such set contains $2^N$ elements (including the identity), and there are $2^N+1$ of such sets~\cite{Lawrence2002}. We will prove that the QCs defined in this way are \pce{} channels with $2^N$ on their diagonal.

First we compute the components of $\mcE^{\text{QC}}_B$ in the basis $2^{-N/2}\left\{\sigma_{\vec \alpha}\right\}$,
\begin{equation}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\frac{1}{2^N}\sum_{i=1}^{2^N} \bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}\bra{\psi_i}\sigma_{\vec l} \ket{\psi_i}.
\end{equation}
To evaluate the components, observe that $\bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}^2=1 \ \ \forall \sigma_{\vec k}\in \text{set}(B)$, and from the formula for the purity of $\ket{\psi_i}$,
\begin{equation}
1=\frac{1}{2^N}\sum_{\vec k} \bra{\psi_i} \sigma_{\vec k} \ket{\psi_i}^2,
\end{equation}
it follows that $\bra{\psi_i} \sigma_{\vec k'} \ket{\psi_i}=0 \ \ \forall \sigma_{\vec k'} \not\in \text{set}(B)$ since there only positive terms in the sum, and $2^N$ of them are already equal to $1$. Therefore,
\begin{align}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}&=1,\nonumber\\ \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l'}&=\left(\mcE^{\text{QC}}_B\right)_{\vec l' \vec l'}=0 \ \ \forall \sigma_{\vec k} \in \text{set}(B) \ \ \forall \sigma_{\vec l'} \not\in \text{set}(B).
\end{align}
To compute $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}$ for $\sigma_{\vec k}, \sigma_{\vec l} \in \text{set}(B)$ with $\vec k \neq \vec l$, observe that $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\left(\mcE^{\text{QC}}_B\right)_{\vec l \vec k}$, \ie{} the matrix corresponding to $\mcE^{\text{QC}}$ is an orthogonal projector. Thus, consider the block,
\begin{equation}
\left[ \begin{array}{cc}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k} & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & \left(\mcE^{\text{QC}}_B\right)_{\vec l \vec l}
\end{array}  \right]=\left[ \begin{array}{cc}
1 & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & 1
\end{array}  \right],
\end{equation}
it is easy to check that the latter is a projector only if $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=0$. Since there are exactly $2^N$ elements of the form $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}$ with $\sigma_{\vec k} \in \text{set}(B)$, the channel $\mcE^{\text{QC}}_B$ is \pce{} with $2^N$ $1$s on its diagonal.