### 2022-02-23 documentos/vector_space.tex
We should get someone from Purdue to do this. They are the boilerplaters.

\subsection{PCE quantum channels as vector spaces} % {{{
% \cpnote{Ojo, cambie el nombre de la seccion. Pendiente de sus comentarios}
% \ddnote{propongo channels as vector spaces como titulo de seccion, pues lo
% actual se entiende mas en el sentido estandar: canales como miembros de un
% espacio vectorial, que tambien es cierto, pero no es el punto aqui}
% \cpnote{Falta lo de las clases de equivalencias, pero aun no encuentro el
% espacio. Quizá cuando haga falta se puede poner}
% \janote{Se me hace que podemos agregar algo al final, ya que es la parte en la que
% se hablan de la regla $2^k$ y regla espejo. No tengo una idea clara, pero si te 
% parece una buena idea podría elaborar más cómo meterlo ahí}
% \par
% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{

In this subsection we will provide a one-to-one relation between PCE quantum
channels and the subspaces of a discrete vector subspace associated with the
indices $\valpha$ labeling the components of a state; see
\eref{eq:N_qubits_rho}. 
Some established facts about vector spaces will allow
us to derive the main features of PCE quantum channels.

Let us start by recalling that the problem of determining complete positivity
of a PCE map can be recast as determining which coefficients $\tau_\valpha$ 
are mapped via $\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to positive eigenvalues $\lambda_\valpha$, as in \eref{eq:lambda:is:A:tau}. 
% . In components, this would read 
% \begin{equation}
% % \left(
% \sum_\vbeta
% \san_{\valpha\vbeta} \,
% % \right)
% \tau_{\vec \beta}=2^Np_{\vec\alpha}
% \label{eq:A:tau:is:p}
% \end{equation}
% where
% $\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
% \sa_{\alpha_N\beta_N}$. 
Using the fact that $a^{-1}=a/4$, so 
% (so $A = A^{-1}/2^N$) 
\begin{equation}
A^{-1} = \frac{1}{4^N} A,
\end{equation}
we can directly invert \eref{eq:lambda:is:A:tau} to obtain 
\begin{equation}
% \tau_{\vec \alpha}= \san_{\vec\alpha\vec\beta} \, p_\vbeta,
% \tau = 
% \san \lambda =2^N \tau
\sum_\vbeta \san_{\valpha \vbeta} \lambda_\vbeta =2^N \tau_\valpha
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. 
This is a remarkable equation, as it provides a method to diagonalize the
Choi matrix of any Pauli diagonal map.

Other two simple but crucial observations are the following. 
For valid quantum channels it holds 
% \begin{equation}
% % \sum_{
% \end{equation}
% $\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
% $\lambda_\valpha=0$ for all $\valpha$ in an arbitrary subset of 
% multi-indices $\Omega$, 
\begin{equation}
\sum_{\valpha \in \Omega} \lambda_\valpha=0 
\quad \implies 
\quad 
\lambda_\valpha=0, \, \forall \valpha \in \Omega
\label{eq:lambda:is:zero}
\end{equation}
for an arbitrary subset of multi-indices $\Omega$, 
as each member of the sum is greater than or equal to zero, 
due to complete positivity of the underlying channel. 
Finally, setting $\valpha =0$ in \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha)=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
% If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
If we now assume that $\tau_{\vec\alpha}=1$, and calculate the difference between
\eref{eq:sum:alpha:is:2N} and
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$
(which follows from \eref{eq:tau:is:A:p} and $\tau_{\vec\alpha}=1$), one obtains
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three
multi-indices, $\valpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$.

Let us now work out the nature of the aforementioned connection.  Let the
components of  $\vbeta$ be such that $\beta_l = \beta \delta_{lk}$ for a
particular particle index $k$, with $\beta$ such that $\sa_{\alpha_k\beta}=1$.
Since $\sa_{\alpha 0}=1$ for any $\alpha$, this particular choice of $\vbeta$
belongs to  $\Phi(\vec\alpha)$, so that \eref{eq:coneccion:dos:indices:A}
indeed holds. However, for such election of $\vbeta$ the above relation reduces
to 
\begin{equation}
\sa_{\beta\gamma_k}=\sa_{\beta\gamma_k^\prime}
\label{eq:14}
\end{equation}
for all $\beta$ such that $\sa_{{\alpha_k}\beta}=1$.  One can verify, by
working out the different cases, that \eref{eq:14} is equivalently expressed as 
\begin{equation}
\gamma_k^\prime=\alpha_k\oplus{}\gamma_k
\label{eq:oplus}
\end{equation}
where the $\oplus{}$ denotes the operation of the Klein group; see Figure
\ref{tab:1} for a detailed description. 
\par
% 
% 
% % Since all elements of $A$ are $\pm1$, and using that $A_{\valpha \vbeta}=1$
% % of $\vbeta \in \Phi(\vec\alpha)$, we may rewrite \eref{eq:coneccion:dos:indices:A}
% % as
% % \begin{equation}
% % \san_{\vbeta\vec\gamma} \san_{\vbeta\vec{\gamma^\prime}} 
% % = 
% % 1
% % , \quad
% % \forall \vec\beta\in\Phi(\vec\alpha),
% % \label{eq:12}
% % \end{equation}
% % which can be further decomposed in 
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}
% %   \ldots
% % %   \cdot\ldots\cdot 
% % \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% % \label{eq:13}
% % \end{equation}
% % Moreover, since 
% % 
% % 
% % 
% % Since $\sa_{\alpha\beta}=\pm1$, we may rewrite (\ref{eq:11}) as
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% % \label{eq:12}
% % \end{equation}
% % This must hold for all $\vec\beta\in\Phi(\vec\alpha)$. From this follows that
% % (\ref{eq:12}) may be rewritten as
% % \begin{equation}
% % \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
% % \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}.
% % \label{eq:13}
% % \end{equation}
% % Here we show that (\ref{eq:13}) implies (\ref{eq:14}). 
% Since \eref{eq:13} holds for 
% all $\vec\beta\in\Phi(\vec\alpha)$, we may choose a specific such $\vec\beta$, namely
% $\beta_l=0$ for all $l\neq k$ and $\beta_k=\beta$, where $\beta$ is an arbitrary number
% such that
% $\sa_{\alpha_k\beta}=1$. 
% % \begin{equation}
% % \sa_{\alpha_k\beta}=1.
% % \end{equation}
% This $\vec\beta$, as is readily verified, belongs to $\Phi(\vec\alpha)$, so
% that \eref{eq:13} indeed holds for this $\vec\beta$. On the other hand, it is
% then readily verified that, for this $\vec\beta$, \eref{eq:13} reduces to 
% % (\ref{eq:14}), whcih therefore follows. 
% % \flnote{Leve modificaci\'on de las ecuaciones anteriores, y adecuaci\'on del
% % texto siguiente. } \cpnote{francois, recuerdo 
% % que esto lo discutimos. No se si en el momento me convencí, pero ahora no veo 
% % proque se tiene que cumplir para los indices individulaes}. 
% % We now show in Appendix
% % \ref{app:proof} that from this follows that
% % for all $1\leq k\leq N$, $\gamma_k$
% % and $\gamma_k^\prime$ are connected by
% \begin{equation}
% \sa_{\beta\gamma_k}\sa_{\beta\gamma_k^\prime}=1
% \label{eq:14}
% \end{equation}
% for all $\beta$ such that $\sa_{{\alpha_k}\beta}=1$. 
% One can verify, by working out the different cases, 
% that \eref{eq:14} is equivalently expressed as 
% \begin{equation}
% \gamma_k^\prime=\alpha_k\oplus{}\gamma_k
% \label{eq:oplus}
% \end{equation}
% where the $\oplus{}$ denotes the operation of the Klein group; see Figure
% \ref{tab:1} for a detailed description. 
% \par
% }}}
% Espacio vectorial {{{

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$.
We notice that the Klein group described in \fref{tab:1} is actually isomorphic
to the two dimensional vector space over the field of two elements $\{0,1\}$.
Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, and say that for quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preserved, and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

\begin{figure} % {{{
\begin{center}
\begin{tabular}{| c | c c c c |}
\hline
$\oplus{}$ & 0 & 1& 2 & 3\\
\hline
0 & 0 & 1 & 2 & 3\\
1 & 1 & 0 & 3 & 2\\
2 & 2 & 3 & 0 & 1 \\
3 &3 & 2 & 1 & 0\\
\hline
\end{tabular}
\caption{Table for the $\oplus{}$ operation defined in (\ref{eq:14}). Note that
the operation is an abelian group, in fact it corresponds to the 
Klein group, where the neutral element is $0$. This is the reason for choosing
an additive notation for the operation defined in (\ref{eq:14}). }
\label{tab:1}
\end{center}
\end{figure} % }}}

% % Lo que quiero quitar {{{
% \cpnote{Propongo quitar desde aca}
% Note further that if we perform the identification
% \begin{subequations}
% \begin{eqnarray}
% 0&\to&(0,0)\label{eq:identa}\\
% 1&\to&(0,1)\label{eq:identb}\\
% 2&\to&(1,0)\label{eq:identc}\\
% 3&\to&(1,1)\label{eq:identd}
% \end{eqnarray}
% \label{eq:ident}
% \end{subequations}
% then the $\oplus$ operation reduces to simple vector addition in
% binary arithmetic. If we therefore denote by $\overline{\alpha}$
% \cpnote{I dont think this is necesary. specially if we will write something
% else}
% the multi-index $\vec\alpha$ converted to a binary vector by the componentwise
% application of (\ref{eq:ident}),
% the connection between $\vec\alpha$, $\vec\gamma$, and $\vec{\gamma^\prime}$ defined by (\ref{eq:11}) is rewritten 
% as
% \begin{equation}
% \overline\gamma=\overline\alpha+\overline{\gamma^\prime}
% \label{eq:17}
% \end{equation}
% where the addition sign here refers to ordinary binary vector addition. 
% \cpnote{Hasta aca}
% % }}}

From this readily follows an amusing property: the set of all multi-indices
$\vec\gamma$ for which $\tau_{\vec\gamma}=1$, is closed under binary
vector addition, in other words, it forms a vector subspace of the set
of all multi-indices.
A moment's consideration will further show that the above reasoning can be
inverted, that is, that if we set all $\tau_{\vec\gamma}$ equal to one
whenever the $\vec\gamma$ belongs to a given vector subspace of the set of
all indices, then $\tau$ indeed has an image
which has
only positive components. 
In other words, there is a one-to-one correspondence between 
a quantum channel, and a vector subspace of the aforementioned space. 
% \cpnote{Aca quizá que Jose Alfredo ponga un ejemplo de una 
% figura en la que se tenga algo que es un canal y algo que no es  un canal. 
% Quizá también se puede mencionar qeu el subespacio mas chico es el canal totalmente 
% depolarizante y el subespacio mas grande es la identidad} 
\par

% }}}
% Generacion de soluciones {{{
With this information, we present a procedure to generate all solutions: we start out
from the solution having $\tau_{\vec0}=1$, with everything else $0$. We may
then successively switch $\tau_{\vec\alpha}$'s
% \janote{agregué el subíndice $\vec\alpha$} 
to $1$ for various values of $\valpha$,
taking care immediately to set equal to one the components of $\tau$
% \janote{puse la $\tau$ como vector y hay otro antes}
that
correspond to values of $\vbeta$ generated by the previously switched values of
$\valpha$ via the operation $\oplus$. Doing so, in an ordered way, allows one
to generate
all PCE quantum channels with a given set of preserved components, without the need
of exploring the exponentially large space of all PCE maps.  \par 
% }}}}

\begin{figure} % {{{
\begin{center}
% Del dcumento francois.pdf pg 8 agarramos:
% e11, e4 y E7
\begin{tikzpicture}[x=0.55cm, y=0.55cm] % {{{
\pgfmathsetmacro{\unitstep}{5.2}
\begin{scope}[shift={(1*\unitstep,0)}] % Depolirizng channel {{{
\node at (-0.5,0.5) {(a)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,0)}] % Interseccion {{{
\node at (-0.5,0.5) {(b)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(3*\unitstep,0)}] % E4 {{{
\node at (-0.5,0.5) {(c)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(1*\unitstep,-\unitstep)}] % E7 {{{
\node at (-0.5,0.5) {(d)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(2*\unitstep,-\unitstep)}] % E11 {{{
\node at (-0.5,0.5) {(e)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\begin{scope}[shift={(3*\unitstep,-\unitstep)}] % Identity channel {{{
\node at (-0.5,0.5) {(f)} ;
\cuadritotikz{}
\begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
\end{scope} % }}}
\end{tikzpicture} % }}}
\end{center}
\caption{
% \cpnote{Revisar, solo tiene una iteracion por parte mia}
% Examples of diagrams for several 2-qubit PCE quantum channels: (a) the totally
% depolarizing channel, (b) a PCE channel that preserves four components.
% (c), (d) and (e) show three generators \cpnote{Quizá poner los nombres de los
% generadores acá}; the combination of any two of them yields the channel in (b).
% (f) represents the identity map. 
% 
% \cpnote{Mejorar el caption cuando me ponga de acuerdo con la gente}
% Aca poner 6 PCEs. (f) Identidad, (a) totally depolarizing, (c, d y e) tres
% generadores, (b) un canal que resulte del generador. Sobra un espacio para
% ilustrar alguna otra cosa. Quiza poner otro generador para que se vea que no
% hay una manera unica de generar esos bichos.
Examples of diagrams for several 2-qubit PCE quantum channels: (a) the totally
depolarizing channel, (b) a PCE channel that preserves four components, the
normalization component, one local component of qubit 2 and two correlations between the 
two qubits.
(c), (d) and (e) show the three generators $\mcG_{(1,0)}$, $\mcG_{(3,2)}$ and
$\mcG_{(2,2)}$, respectively; the combination (overlap of diagrams) of any two
of them yields the channel in (b).
(f) represents the identity map. 
}
\label{fig:examplesPCE}
\end{figure} % }}}
% Los canales tiene 2^K componente {{{

% **********************************
% 
% We can show that all PCE quantum channels preserve $2^K$ components. 
% In
% fact, with the above picture, it is clearly revealed that the vector space has
% dimension $2N$. Moreover, 
% consider any subspace and chose the maximum number of linearly independent
% elements (say $K$).  Any of the  $2^K$ subsets of those elements will yield a
% unique element. Moreover, since $\valpha + \valpha = \vec 0$, these are all
% elements of the vector space. This means that any PCE quantum channel will preserve $2^K$ 
% components, for $K$ an integer. See \fref{fig:examplesPCE} for some examples. 
% \par 
% 
% **********************************
% **********************************

We can show that all PCE quantum channels preserve $2^K$ components. 
First recall that a vector space of dimension $d$ over a field of $q$ elements
has $q^d$ elements~\cite{Roman2008}.
Now $V$ is a vector space on a field of two elements having dimension $2N$. 
% The dimension of the vector space $V$ of all multi-indices is $2N$, and 
We have
seen earlier that
\begin{equation}
W=\left\{
\vec\alpha:\vec\alpha\in V, \tau_{\vec\alpha}=1
\right\}
\label{eq:19}
\end{equation}
is a subspace of $V$. 
As such, $W$ has a given dimension $K$, which means that
$W$ has $2^K$ elements. In other words, a set of indices $\tau_\valpha$ with the property
discussed above can only have $2^K$ elements equal to 1, for a given integer
$K$. \par

% **********************************
% **********************************
% 
% \cpnote{Dejar que la raza elija uno de los dos parrafos anteriores}\janote{El
% primero (simón, cambié de opinión)} \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCE quantum channels exist that preserve  $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}\label{eq:conteo:simmetry}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest a relation  between individual channels that
preserve $K$ and $2N-K$ Pauli components that for the time being
has escaped our efforts to identify. 
\par % }}}

% }}}