### 2022-03-09 documentos/a_and_generators.tex
Sobre simetrías de A y los generadores

\documentclass[11pt,dvipsnames]{article} % {{{

\usepackage{geometry}
\geometry{total={170mm,240mm}, left=20mm, top=20mm}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{enumerate} 
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{physics}

\usepackage{bbold}

% To add graphics
\usepackage{graphicx}
% Path to figures
\graphicspath{ {../images/} }

\usepackage[draft,inline,nomargin]{fixme} \fxsetup{theme=color}
\definecolor{jacolor}{RGB}{200,40,0} \FXRegisterAuthor{ja}{aja}{\color{jacolor}JA}

%% commands
\newcommand{\valpha}{\vec\alpha}
\newcommand{\pceg}{\mathcal{G}_{\valpha}}
\newcommand{\1}{\mathbb{1}}

\author{José Alfredo de León}
\title{a and generators}
\begin{document}
\maketitle

One can express $\vec \tau(\pceg)$, the vector $\vec\tau$
of a PCE generator $\pceg$, in terms of $a(\alpha_{j_k})$, the rows or columns
of matrix 
\begin{align}
a=
\left(
\begin{array}{cccc}
 1 & 1 & 1 & 1 \\
 1 & 1 & -1 & -1 \\
 1 & -1 & 1 & -1 \\
 1 & -1 & -1 & 1 \\
\end{array}
\right).
\end{align}
The aforementioned expression is 
\begin{align}
\vec \tau(\pceg)=\frac{a(\alpha_{j_1})\otimes\ldots\otimes a(\alpha_{j_N})+\1_{4^N}}{2},
\end{align}
where $\1_{4^N}$ is a $4^N$-length vector with all its components equal to 1.

Note that the rows or columns $a(\alpha_{j_k})$ of matrix $a$ 
are eigenvectors of the reflection matrix
\begin{align}
R=
\left(
\begin{array}{cccc}
 0 & 0 & 0 & 1 \\
 0 & 0 & 1 & 0 \\
 0 & 1 & 0 & 0 \\
 1 & 0 & 0 & 0 \\
\end{array}
\right)
\end{align}
with eigenvalues $r_{0,1,2,3}=1,-1,-1,1$. In other words, we can identify
the `symmetric' (`anti-symmetric') eigenvectors as those with eigenvalues
1 (-1).

\janote{probar que (2) si es la tau de un generador}
We shall start proving that $\vec \tau(\pceg)$ has $4^{N-1}$ components 
equal to 1 and the rest 0.

The vector $a(\alpha_{j_1})\otimes\ldots\otimes a(\alpha_{j_N})$ has 
$n_-$  and $n_+$ number of components with minus and plus sign, respectively.
\begin{align}
n_{\pm}&=\sum_k\mathbb{n}_{\pm}\big[a(\alpha_{j_k})\big],
\end{align}
with $\mathbb{n}_{\pm}\big[a(\alpha_{j_k})\big]$ the number of components
with sign $\pm$ in vector $a(\alpha_{j_k})$.\janote{asumiendo que ya 
se probó que el número de +1 es $4^{N-1}$ o $4^{N}$ y de -1 $4^{N-1}$ o 0:}

Now we shall prove the symmetry.
\begin{align}
\Sigma^{(k)}\big[\vec \tau(\pceg)\big]=
\frac{a(\alpha_{j_1})\otimes\ldots\otimes R\big[ a(\alpha_{j_k})\big]\ldots\otimes a(\alpha_{j_N})+\1_{4^N}}{2}
\end{align}

\end{document}