### 2022-04-23 documentos/kraus.tex
Rev_App

\section{PCE channels and decoherence o Physical implementations of PCE maps} % {{{
\label{sec:kraus}
% Intro {{{

% \cpnote{HAcer esta seccion despues de que david ponga lo de decoherencia, 
% para servirla bien}
% \todo{Carlos hace una version}

% {\red [texto viejo introductorio de la seccion] Furthermore, quantum channels enjoy the well-known \textit{Kraus
% representation} (or operator-sum representation) which reads 
% \begin{equation}
% \mcE[\rho]=\sum_i K_i \rho
% K_i^\dagger{},
% \end{equation}
% with $\sum_i K_i^\dagger{}K_i=\one$ (the trace-preserving condition)~\cite{Kraus1983}.
% 
% The Kraus operators of PCE channels provide an insight in the physical processes
% that yield such channels. In this section we investigate and construct such 
% operators and explore the corresponding processes. To study the Kraus operators corresponding to PCE, it is enough to study 
% the ones corresponding to generators, as all PCEs commute and thus to get the 
% Kraus operators of an arbitrary PCE, one can easily combine the Kraus
% operators of its generators. }

% {\red [anterior]
% Given that \pce{} channels are projections, we might ask whether they are fixed
% points of some memory-less (Markovian) process or not. In this section we prove
% that they are, and give two examples of Markovian implementations. The first
% one consist in identifying each \pce{} channel as a fixed point of some pure
% dissipative process. The second, in identifying each \pce{} channel as fixed
% points of some memory-less collision model.}

Lindblad processes arise naturally in many theoretical~\cite{lindblad,kossa,kossa2,Gorini1976,cirac,davalosdivisibility} and
experimental~\cite{PhysRevA.67.042322} settings and are archetypical in decoherence dynamics. Moreover, these processes lead to a monotonically \afnote{monotonic (continuous)} lose of information~\cite{breuerreview} and describe non invertible channels in the asymptotic limit $t\to \infty$ (this can be seen from the monotonically \afnote{(continuous)} decreasing of the determinant, see Ref.~\cite{cirac}).
It is known that not every quantum channel can be seen as a snapshot of a
process
arising from a traditional Lindblad equation or even a time-dependent Lindblad
equation~\cite{cirac,davalosdivisibility}. 
Therefore, an interesting question is whether \pce{} channels can be seen as limit points of
some Markovian processes. In this section we prove what \afnote{that} in fact they are, and
give two examples of Markovian implementations.
%
The first
one consist in identifying each \pce{} channel as a fixed point of some pure
dissipative process. The second, in identifying each \pce{} channel as fixed
points of some memory-less collision model \afnote{The first of them consists in identifying each \pce{} channel as a fixed point of some pure dissipative process, and in the second implementation we relate each \pce{} channel to fixed
points of some memory-less collision model }.

% the most 
% common decoherence 
% 
% with a Markovian master
% equation~\cite{lindblad,kossa,kossa2,Gorini1976},
% \begin{multline}
% \dot \rho(t)=-i [H(t),\rho(t)]+\\
% \sum_k \gamma_k(t)\left( V_k(t)\rho(t)V_k^{\dagger{}}-\frac{1}{2} \left\{V_k^\dagger{} V_k(t),\rho(t) \right\}\right),
% \end{multline}
% where the first part is the so-called Hamiltonian term, the second the dissipative term with $\gamma_k(t)\geq 0$ the relaxation ratios associated with the Lindblad operators $V_k(t)$.
% %
% In the particular case where the Hamiltonian, relaxation rations and Lindblad
% operators are independent of time, we recover the well-known Lindblad master
% equations~\cite{lindblad}. Such equations model time-homogeneous processes,
% implying one-parameter semigroup structures that will be exploited later.
% %
% Besides, these processes lead to a monotonically lose of information~\cite{breuerreview} and describe non invertible channels in the asymptotic limit $t\to \infty$ (this can be seen from the monotonically decreasing of the determinant, see Ref.~\cite{cirac}). Therefore, we might ask whether \pce{} channels can be seen as limit points of some Markovian processes. In this section we prove what in fact they are, and give two examples of Markovian implementations.
% %
% The first
% one consist in identifying each \pce{} channel as a fixed point of some pure
% dissipative process. The second, in identifying each \pce{} channel as fixed
% points of some memory-less collision model.

% }}}
\subsection{Kraus representation} % {{{
To derive the aforementioned implementations, we exploit the existence of the
PCEGs \janote{introducimos esta notación antes?} \afnote{(acabo de colocarla en el 3er parrafo de la sección de generadores)} and their \textit{Kraus representation} (or operator-sum representation)
which, for an arbitrary channel $\mcE$,  reads 
\begin{equation}
\mcE[\rho]=\sum_i K_i \rho
K_i^\dagger{},
\end{equation}
with $\sum_i K_i^\dagger{}K_i=\one$ (the trace-preserving condition)~\cite{Kraus1983}.
% The Kraus operators corresponding to one single qubit PCEs are 
% $\{\openone/\sqrt2, \sigma_i/\sqrt2\}$,
% % \begin{equation}
% % \{\openone/\sqrt2, \sigma_i/\sqrt2\}
% % \label{eq:kraus:PCE}
% % \end{equation}
% corresponding to the operation that leaves the component $\sigma_i$
% invariant~\cite{nielsen_chuang_2011}.
Inspection of the Kraus operators of two qubit PCEGs leads to
the ansatz that the Kraus operators for the generator of 
$\mcG_{\vec \alpha}$ are 
\afnote{By inspection, it is easy to prove that two qubit PCEGs leads to
the ansatz that the Kraus operators for the generator of 
$\mcG_{\vec \alpha}$ are}
\begin{equation}
K_0=\frac{\openone}{\sqrt2}, \quad K_1=\frac{\sigma_{\vec \alpha}}{\sqrt2},
% \left\{\openone/\sqrt2, 
% \left(\bigotimes_{k=1}^N \sigma_{\alpha_k}\right)/\sqrt2\right\}. 
\label{eq:kraus:PCE}
\end{equation}
since the Kraus operators corresponding to one \afnote{a} single qubit PCEs are 
$\{\openone/\sqrt2, \sigma_\alpha/\sqrt2\}$,
corresponding to the operation that leaves the component $\sigma_\alpha$ \janote{$r_0$ and $r_\alpha$}
invariant~\cite{nielsen_chuang_2011}. \ddnote{agrego lo de los flips aqui}{\blue Notice that according to Kraus operators, the generators $\mcG_{\vec \alpha}$ are $N$-qubit flip channels with flip probability of $1/2$, where the joint flip is $\rho \mapsto \sigma_{\vec \alpha} \rho \sigma_{\vec \alpha}$. In fact, tracing out all particles except particle $k$ \afnote{...except the $k-$th one} gives the well known qubit flip channels, \ie{} $\tr_{\not k} \mcG_\valpha=\mcG_{\alpha_k}$ (see~ \eref{eq:generator:individual:qubit}). And more generally, tracing out $m$ particles leaves a $N-m$ particles flip channel \afnote{with weight $1/2$}.}

We shall first show that the Kraus operators in \eref{eq:kraus:PCE} produce \janote{are indeed} a PCE. 
Notice that 
% \begin{equation}
$\sigma_\alpha \sigma_\beta \sigma_\alpha =a_{\alpha \beta} \sigma_\beta$, see~\Eref{eq:1},
% \label{}
% \end{equation}
which in turn imply \afnote{implies} that 
\begin{equation}
\sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha} 
=\san_{\vec \alpha \vec \beta} \sigma_{\vec \beta}.
\label{eq:sigma_property}
\end{equation}
% with $a$ the matrix that appears in \eqref{eq:eigvals-PCE}. 
Next, consider the action of a channel with Kraus representation 
\eqref{eq:kraus:PCE} on a $N$-qubit system:
\begin{align}
\rho \mapsto &\frac{1}{2^N}\sum_{\vec \beta} r_{\vec \beta} \left( \frac{1}{2} \sigma_{\vec \beta} + \frac{1}{2} \sigma_{\vec \alpha} \sigma_{\vec \beta} \sigma_{\vec \alpha} \right)\nonumber \\
=&\frac{1}{2^N}\sum_{\vec \beta} r_{\vec \beta} \frac{1+A_{\vec \alpha \vec \beta}}{ 2}\sigma_{\vec \beta}.
\end{align}
However, since $\san_{\vec \alpha \vec \beta} = \pm 1$, the channel characterized
by \afnote{the} Kraus operators \janote{in} \eref{eq:kraus:PCE} is a PCE channel. Moreover, one can notice that, 
except for the first row, half of the matrix elements of each row is $+1$ and 
half $-1$, which implies that the channel is a PCEG.

% We still have to show that except for $i_1=\cdots=i_n=0$, half of the numbers 
% $(1+\prod_k a^{i_k}_{j_k})/2=1$, or equivalently, half of $\prod_k a^{i_k}_{j_k}=1$. 
% Consider a particular $l$ such that $i_l \ne 0$. This will make half of 
% the $\prod_k a^{i_k}_{j_k}$ 1 and half $-1$ (for fixed $l$). This also implies 
% that half of the correlations of the density matrix are erased and half preserved 
% so the channel is a PCEG.

Observe	 also that a different choice of $\vec \alpha$ in \eref{eq:kraus:PCE}
leads to different channels. This follows from the fact that if two channels were
the same, this would imply that the matrix representation of the corresponding
superoperator of $\rho \mapsto \sigma_{\vec \alpha}\rho \sigma_{\vec \alpha}$
% $\sigma_{\vec \alpha} \cdot \sigma_{\vec \alpha}$ 
% \ddnote{en este punto me confundio bastante esta notacion, contrapropongo $\rho
% \mapsto \sigma_{\vec \alpha}\rho \sigma_{\vec \alpha}$, que es una notación que
% hemos venido usando.}
would have to be the same, which is clearly false. Since there are $4^n$ different 
$\vec \alpha$, this implies that all PCEG (plus the identity map) are in one to 
one correspondence.

% \cpnote{We might have spoken about the notation of the generators and here we 
% can add a sentence about it, but I will wait till JA writes his part.}
% Two remarks. (i) The notation that we develop with Jose Alfredo coincides with the 
% generator. This notation is based on the action on each individual qubit. (ii) Counting the
% number of generators here is trivial. It is simply $4^n-1$ as previously noticed 
% from other arguments. 
% }}}
\subsection{Pure dissipative implementation} % {{{
% {\red [viejo] Any PCE channel can be seen as the fixed point of some
% decoherence process\cpnote{esto suena obvio, pero no lo es. Es decir no todos
% los canales son resultado de lindblad.}
In this section we show that any PCE channel can be seen as the
fixed point of some decoherence process, starting with PCEGs and 
then generalizing \afnote{extending} to more general channels. 
% Me gustaria modificar esta
% frase para que pueda ser apreciada mejor} \ddnote{la modificacion fue leve
% debido a que la mayor parte del contenido esta en la introduccion de la
% seccion}. 
% We start with PCEGs. 
Consider the following dynamical process that
implements $\mcG_{\vec \alpha}$ when $t\to \infty$,
\begin{align}
\mcG_{\vec \alpha,t}[\rho]&=e^{- \gamma t} \rho + (1-e^{-\gamma t})\mcG_{\vec \alpha}[\rho]\nonumber \\
&=\frac{\left( 1+e^{- \gamma t} \right)}{2} \rho +\frac{\left( 1-e^{- \gamma t} \right)}{2} \sigma_{\vec \alpha} \rho \sigma_{\vec \alpha},
\end{align}
where $\gamma>0$. It is easy to show that the family of channels $\mcG_{\vec
\alpha, t}$ parametrized with $t \geq 0$ forms a one-parametric semigroup,
\ie{} $\mcG_{\vec \alpha, t_1}\mcG_{\vec \alpha, t_2}=\mcG_{\vec \alpha,
t_1+t_2}$, therefore $\mcG_{\vec \alpha, t}$ describes a dissipative
time-homogeneous Markovian process, those are always characterized by some
Lindblad generator~\cite{lindblad} \afnote{Me parece que se podría conectar mejor esta última frase (pero no pude XD)}. The Lindblad generator of $\mcG_{\vec
\alpha, t}$, denoted by $\mcL_{\vec \alpha}$, can be obtained with \afnote{by} the standard
procedure,  
\begin{equation}
\mcL_{\vec \alpha}[\rho]=\left. \frac{d\mcG_{\vec \alpha,t}[\rho]}{dt}\right|_{t=0}=\frac{\gamma\left(\sigma_{\vec \alpha}\rho \sigma_{\vec \alpha} -\rho \right)}{2},
\end{equation}
where the unique Lindblad operator associated with the relaxation ratio
$\gamma/2$ is simply $\sigma_{\vec \alpha}$. Notice that $\sigma_{\vec \alpha}$
is trace-less, therefore the process is pure dissipative~\cite{cirac}.

Since PCEGs commute, we can describe easily any other PCE channel as a fixed
point of a decoherence process. For them, the Lindblad generators are the sum
of the Lindbladians of the corresponding generators. As an example, consider
the channel depicted in \fref{fig:two:qubit:examples} b), it is equal to
$\mcG_{(0,3)}\mcG_{(3,3)}$, therefore it is the fixed point of the dissipation
process described with the following Lindbladian,
\begin{equation}
\mcL[\rho]=\frac{\gamma_{(0,3)}\left(\sigma_{(0,3)}\rho\sigma_{(0,3)}-\rho \right)}{2}
   +\frac{\gamma_{(3,3)}\left(\sigma_{(3,3)}\rho\sigma_{(3,3)}-\rho
\right)}{2},
\end{equation}
where $\gamma_{(0,3)}$ and $\gamma_{(3,3)}$ are positive and correspond to the Lindblad operators $\sigma_{(0,3)}$ and $\sigma_{(3,3)}$. Notice that such election of Lindblad operators is not unique, as the PCE channel described here is also equal to $\mcG_{(0,3)}\mcG_{(3,0)}$.

% }}}
\subsection{Collision model implementation} % {{{
We show now that PCE channels can also be implemented with \textit{simple
collision models}~\cite{Ziman2010}. To do this, observe that using \afnote{by employing} the Stinespring dilation theorem~\cite{Stinespring2006}, PCEGs can
be implemented using an unitary over the system and an
ancilla. Since PCEG
always have Kraus rank $2$, one can
always choose a qubit as the ancillary system. Concretely,
\begin{equation}
\mcG_{\vec \alpha}[\rho]=
    \tr_{\text{qubit}} \left( 
         U_{\vec \alpha} \left(\rho\otimes \projj{0}\right) U_{\vec \alpha}^\dagger{} 
     \right),
\end{equation}
where $\tr_{\text{qubit}}$ denotes partial trace over the ancillary qubit, with
the unitary defined as follows,
\begin{align}
U_{\vec\alpha}\ket{\psi} \ket{0}= \frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}+\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right),\\
U_{\vec\alpha}\ket{\psi} \ket{1}=\frac{1}{\sqrt{2}} 
   \left( \ket{\psi} \ket{0}-\sigma_{\vec\alpha} \ket{\psi}\ket{1} \right).
\label{eq:Nqubit_stinespring}
\end{align}
Therefore, any concatenation of PCEGs can be described as a collision model
with as many collisions as generators needed. In fact, generators are described
with one collision. For the general case consider some PCE channel $\mcE$
generated with $\left\{ \mcG_{\vec \alpha_1}, \mcG_{\vec \alpha_2}, \dots,
\mcG_{\vec \alpha_M} \right\}$. For this we can define an environment
consisting of \afnote{on} $M$ qubits initially in the state $\left( \projj{0}
\right)^{\otimes M}$, or equivalently one qubit with the additional assumption
that its state is reset to $\ket{0}$ after each collision (memory-less
collisions). The collision with the $k\text{th}$ \afnote{$k\text{-th}$} particle is described by
$U_{\vec \alpha_k}$, which acts solely over the system and the $k\text{th}$
particle. Therefore $\mcE$ can be written as follows,
\begin{equation}
\mcE[\rho]=\tr_{\text{E}} \left[ \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M}\right) \rho \otimes \left(\projj{0}\right)^{\otimes M} \left(U_{\vec \alpha_1} \dots U_{\vec \alpha_M} \right)^\dagger{} \right],
\end{equation}
where $\tr_{\text{E}}$ is the partial trace over the total of the ancillary qubits \afnote{over the whole ancillary qubits}. Notice that as PCEGs commute, the order of the collisions can be any \afnote{the order of the collisions is irrelevant}.
% \begin{itemize}
% \item Relate to decoherence, for PCEG\ddnote{hecho}
% \item Play with non PCEG PCEs\ddnote{hecho}
% \item Dependiendo de si sale algo bonito, podemos poner algun modelo en
% particular para todos los canales (sistema central + ambiente). Seria de buscar
% dilataciones sencillas y razonables por lo menos a los canales generadores.\ddnote{hecho, es el asunto de collision models}
% \todo{Davalos lidera la discusión de ese punto, pero no escribe nada}
% \end{itemize} 
% 
% \begin{equation}
% \mcL_{\vec \alpha}[\Delta]:= \frac{\gamma}{2}\left( \sigma_{\vec\alpha} \Delta \sigma_{\vec\alpha} -\Delta \right).
% \label{eq:Nqubit_semigroup}
% \end{equation}

% }}}
% }}}