### 2022-05-03 documentos/intro.tex
Is there an award for this?

\section{Introduction} % {{{

Quantum correlations~\cite{PhysicsPhysiqueFizika.1.195,
PhysRevLett.126.170404,Ollivier2001}, including entanglement~\cite{Horodecki},
are an important resource for a wide variety of tasks that
include teleportation~\cite{PhysRevLett.70.1895}, quantum
computation~\cite{nielsen_chuang_2011}, and others~\cite{Sapienza2019}.
However, this resource is also extremely delicate~\cite{nielsen_chuang_2011,Schlosshauer2004}, specially 
for multi-particle systems~\cite{Horodecki},  and important part
of the efforts of the community implementing quantum technologies is 
devoted to tackle this issue from an experimental \cite{Mooney_2021,Briegel1998} and
theoretical \cite{Bennett1996,Georgescu2020,RevModPhys.87.307} point of view. 
The process by which quantum correlations are unintentionally dissipated
is called decoherence~\cite{Schlosshauer2004,sep-qm-decoherence}.
One of the main tools to study the effects of decoherence are 
quantum channels. Quantum channels can describe quantum
noise~\cite{davalosdivisibility,ruskai}, open quantum systems
dynamics~\cite{nielsen_chuang_2011,breuerbook}, and recently even coarse
graining~\cite{Duarte2017,pinedacg}. 
%
% The identification of families of quantum channels among the whole convex set
% of quantum channels associated to a Hilbert space, is crucial to understand
% quantum noise~\cite{davalosdivisibility,ruskai}, open quantum systems
% dynamics~\cite{nielsen_chuang_2011,breuerbook}, and recently even coarse
% graining~\cite{Duarte2017,pinedacg}. 
One of the main difficulties to characterize quantum channels is that,
similar to quantum states, the parameters needed to describe them
increases overwhelmingly rapidly with the dimension of the Hilbert space.
%
Moreover, such parameters are constrained in a complicated way by 
physical conditions, such as complete positivity ~\cite{zimansbook}.
% Such parameters are not fully independent, the complete positivity condition imposes the regions that have physical meaning~\cite{zimansbook}.
% 
Describing in detail families of channels having a given property provides
insight into the jungle of quantum operations.
For the qubit case there are several studies concerning the unital case, for which
non-trivial properties can be described using only $3$ parameters, which in
turn form the well known tetrahedron of Pauli
channels~\cite{Rybar2012,ruskai,davalosdivisibility}. 
% For general
% finite-dimensional channels, several families of channels have been defined.
More generally, in~Ref.~\cite{nathanson2007pauli} the authors study families of 
convex combinations of quantum-classical channels that relate to
% have a close relation with 
unital qubit channels with positive eigenvalues, and give a
generalization of the Bloch sphere. 
% {\blue We will show later that some of the channels studied here are also
% quantum-classical.}\ddnote{Esta frase no me gusta, pero no se me occure nada
% mejor en tanto a la escritura. Propongo omitirla, despues se volvera a
% mencionar esto pero en donde corresponde.}
Similarly, a generalization of Pauli channels based on mutually unbiased
measurements is introduced and studied in Ref.~\cite{Siudzinska2020}. Other
studies of channels beyond the qubit can be
found~\cite{pub.1117408526,pub.1060516327,pub.1014660854}.
% \cpnote{Me gusta la idea de ir sobre un par de familias, pero no veo para que 
% mencionar a las MUBs aca, ni paraque introducir la sigla. Ojala podamos 
% discutir que familias mencionamos acá}\ddnote{atendido como fue acordado en la reunion :)}

\par
In this work we present a generalization of idempotent Pauli channels, \ie{}
the qubit flip operations (bit, phase and bit-phase when the flip probability is $1/2$), total depolarizing qubit
channel and the identity channel, to the case of $N$ qubits. 
% \cpnote{Mencionar decoherencia} 
The generalization is done by extending
Pauli observables to \textit{Pauli strings} (tensor products of Pauli
matrices)~\cite{KIMURA2003339,Lawrence2002}, the resulting maps are unital and
diagonal in the Pauli strings' basis.
In addition, these channels describe the decoherence process in a similar way
as the aforementioned
ones for the single-qubit case.
From now on we refer to these maps as \textit{Pauli component erasing} (\pce{})
maps. We
derive necessary and sufficient conditions for such mappings to be completely
positive trace preserving, thus, showing which idempotent operations are valid
quantum channels (\pce{} channels). Additionally, we characterize the algebraic
properties of the set of \pce{} channels, showing that it is equipped
with a semigroup structure, which is later used to derive physical
implementations.
%
\par
The paper is organized as follows, in section~\ref{sec:pce_maps} we recall
the properties of quantum channels needed to proceed with the definition 
of 
% the tools to unders
% refreshing of quantum channels theory and the definition of 
\pce{} maps. In
section~\ref{sec:math} we diagonalize analytically the Choi matrix for
arbitrary \pce{} maps and characterize completely positive maps by interpreting
\pce{} maps as finite vector spaces. 
We study the
generators of the semigroup structure associated to the set of \pce{} channels in section~\ref{sec:generators}, and we use
them to derive meaningful physical interpretations of \pce{} channels in
section~\ref{sec:kraus}, as well Kraus operators of the generators.
% 
To finish, we conclude in section~\ref{sec:conclusions}.
% }}}