### 2022-05-03 documentos/generators.tex
multiindex

\section{Generators} % {{{
\label{sec:generators}
%\todo{Francois escribe una primera version del primer punto, y luego quizá Jose
%Alfredo o Carlos la terminan}

We now discuss the existence of a generator set for all PCE quantum channels, 
how to label each of them uniquely as $\pceg$ according to its local action 
on every qubit in the system, and, finally, we will discuss a symmetry of
both PCE quantum channel generators and $A$, the matrix that diagonalizes
the Choi matrix $\mcD_N$ of a PCE map.

There exists a subset of PCE quantum channels that generates the entire set; 
the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCE channels,
this implies that all PCE quantum channels can be obtained as
compositions of PCE channels corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCE quantum channels that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
PCE generators and the composition of any two of them yield the PCE channel
corresponding to (b).

PCE generators \afnote{(hereafter PCEGs)} may be characterized by its local action on every qubit 
in the system. This action can be encoded using a multi-index $\vec\alpha$, 
as in \eqref{eq:pauli_strings}, hence each of the different $4^N$ 
vector indices may be uniquely related to each of the PCE generators
and thus denoted as $\mcG_{\vec\alpha}$.
% \cpnote{
The proof is simplified if one uses the Kraus representation developed in 
\sref{sec:kraus}, so we postpone the demonstration to appendix \ref{sec:2_qubits}.
\janote{referir a fig 5} For single qubits, the identity corresponds to $\mcG_0$, shown in 
\fref{fig:one:qubit:examples}(b), whereas $\mcG_3$ is shown in in 
\fref{fig:one:qubit:examples}(c).
% }
The 2-qubits PCE generator represented in subfigure (c)
of \Fref{fig:examplesPCE} acts on the first qubit (first column) as mapping its
Bloch sphere to de $x$-axis, and on the second qubit (first row) acts an identity,
hence it is labeled $\mcG_{(1,0)}$.
See \Fref{fig:appex:twoQ_PCEGs} for the notation of all 2-qubit PCE generators.

A reflection symmetry is identified for PCE generators.
Consider the map $\Sigma^{(k)}$ that reflects an multi-index $\valpha$ with 
respect to the $k$-th axis. This map leaves all components of $\valpha$
invariant, except the $k$-th component, which is transformed according to 
\begin{equation}
% \Sigma: 
0\mapsto3,\quad3\mapsto0,\quad1\mapsto2,\quad2\mapsto1. 
\label{eq:Sigma}
\end{equation}
% and the associated 
The maps have the following properties:
\begin{enumerate}
\item $\Sigma^{(k)}(\vec\alpha)\oplus \Sigma^{(k)}(\vec\beta)=\vec\alpha\oplus \vec\beta$, and
\item $\Sigma^{(k)}(\vec\alpha)\neq\vec\alpha$.
\end{enumerate}
From the first property, we now obtain
\begin{equation}
\valpha=\Sigma^{(k)}(\valpha) \oplus \Sigma^{(k)}(\vec0). 
\label{eq:Sigma:oplus}
\end{equation}
This implies that if $\Sigma^{(k)}(\vec0)$ belongs
to a channel then 
$\tau_\valpha = \tau_{\Sigma^{(k)}(\vec\alpha)}$, where we used 
the fact that for channels, the non-zero elements are closed
under $\oplus$. In other words, the components of a PCE channel
are symmetric under reflection over the $k$-th axis. Now consider the case in which 
 $\Sigma^{(k)}(\vec0)$ does not belong to generator. Then 
$\tau_\valpha \ne \tau_{\Sigma^{(k)}(\vec\alpha)}$, since
the case $\tau_\valpha = \tau_{\Sigma^{(k)}(\vec\alpha)}=1$
is forbidden due to \eref{eq:Sigma} and the case 
$\tau_\valpha = \tau_{\Sigma^{(k)}(\vec\alpha)}=0$ is also forbidden 
because for generators, the codimension of the associated vector space 
is 1. This means that the components of a PCE channel are antisymmetric
with respect to reflection over the $k$-th axis. 
Indeed, the 2-qubit PCE generators 
$\mcG_{(1,0)}$, $\mcG_{(3,2)}$ and $\mcG_{(2,2)}$
represented in subfigures (c), (d) and (e) of \Fref{fig:two:qubit:examples}, 
respectively, are either symmetric or anti-symmetric under reflection with 
respect to lines that divide the diagram in half vertically and horizontally.

\cpnote{Revisar este párrfo. Solo ha pasado una persona por ahi}
% 
% 
% The symmetry for a PCE generator $\pceg$ is encoded in the 
% $N$-dimensional array of all $\taus$. The Cartesian grid associated with it 
% is either symmetric or anti-symmetric under reflection of the 
% indices in the $k$-th position in $\vec\alpha$. 
% Consider the map
% \begin{equation}
% % \Sigma: 
% 0\mapsto3,\quad3\mapsto0,\quad1\mapsto2,\quad2\mapsto1
% \label{eq:Sigma}
% \end{equation}
% Indeed, any generator is a subspace containing half the elements of the full vector space.  
% Map \eref{eq:Sigma} applied to the $k$-th qubit has the following properties
% \begin{enumerate}
% \item $\Sigma(\vec\alpha)\oplus \Sigma(\vec\beta)=\vec\alpha\oplus \vec\beta$.
% \item $\Sigma(\vec\alpha)\neq\vec\alpha$
% \end{enumerate}
% From the first property, we now obtain
% \begin{equation}
% \vec\alpha=\Sigma(\vec\alpha) \oplus \Sigma(\vec0)
% \end{equation}
% We now see that, depending on whether $\Sigma(\vec0)$ belongs (respectively does not belong) to the generator,
% $\vec\alpha$ belongs to the generator if and only if $\Sigma(\vec\alpha)$ belongs (respectively does not belong) to the generator.  
%

% \janote{Aquí iría la prueva de la simetría de reflexión que está en el horno aún.}
% For instance, see that the 2-qubit PCE generators 
% $\mcG_{(1,0)}$, $\mcG_{(3,2)}$ and $\mcG_{(2,2)}$
% represented in subfigures (c), (d) and (e) of \Fref{fig:two:qubit:examples}, 
% respectively, are either symmetric or anti-symmetric under reflection with 
% respect to lines that divide the diagram in half vertically and horizontally. 
% The same symmetry is identified for the rows and columns of $A$ if one 
% labels each position in a row or column $A^{(i)}$ as $A^{(i)}_{\vec\alpha}$.
% \janote{Aquí habría que agregar la prueva de simetría para la $A$.}

\begin{figure} % {{{
\centering
\includegraphics[width=\columnwidth]{oneQ_PCEGs-crop}
\caption{Connection between rows and columns of $a$ and single qubit PCE generators.
The rows or columns of $a$ and the diagrams of the generators are either 
symmetric or anti-symmetric under reflection.
\janote{Propuesta de figura para mostrar que generadores 
y la a obedecen las mismas simetrías.}
}
\label{fig:A_and_PCEGs_connection}
\end{figure} % }}}

% \ddnote{Tengo entendido que esto se queria probar usando las simetrias, esto ya
% esta probado en la secc V}
It is worth pointing out that $\san$ (and thus $a$) [see \eref{eq:san}] 
encodes all the information of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
From the tensor power of matrix $a$ one can infer the components $\{\taus\}$ of a PCE
generator $\pceg$ by taking row (or column) $\alpha$ of $A$ and replacing
$-1$ with $0$. 
The proof is given in \appref{sec:2_qubits} and in \fref{fig:A_and_PCEGs_connection} we illustrate
this connection for the single qubit case.

% \cpnote{Sienot que no tiene flujo. Quizá se le pueda dar flujo usando el final 
% del primer parrafo}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% 
% The set of all PCE quantum channels may be sorted in subsets, 
% which we will call equivalence classes, whose
% elements are equivalent via particle swaps and local-basis element
% permutations. 
% \janote{El párrafo que pretendía ser este tendrá que ir en otra parte. 
% Aún hay que decidirlo con los demás.}

% }}}