### 2022-08-05 documentos/answer_referees_PRA.txt
My bad

Dear Editor,

We are happy to receive positive feedback from the referees. The points
raised by them are answered in what follows.

Answer to the First Referee  {{{

*********************************************************************
1, Is there any difference or they are interchangeable between Pauli
diagonal operation and Pauli diagonal map? Please define them clearly
in the text when they first appear.

Answer:
There is no difference. We have replaced all occurrences
of "PCE operation" with "PCE map" (or similar), see list of changes.

*********************************************************************
2, It is clear that the channel Eq. (12) defines a super operator
acting on any B(H), but for the Choi matrix in Eq. (13), why it is a
super-operator?

Answer:
We have changed the wording. We now call D_N a "matrix"
(as it is a Choi matrix) instead of a superoperator.

*********************************************************************
3, I find this sentence “Let the components of \beta be such that
\beta_l =\beta\delta_lk for a particular particle index k, with \beta
such that a_\alpha_k\beta  = 1” rather confusing.

Answer:
We agree, the statement was confusing. We are now providing a 
new (and hopefully better) explanation.

*********************************************************************
4, Although two mathematical decoherence models are given, the
physical motivation of studying this general many qubit PCE channel is
not quite clear? It also would be better to motivate it in the
beginning.

Answer:
This paper grew from the sheer curiosity of why, for a single qubit, some
projections are allowed and others not. After some time, the beautiful 
mathematical structure presented in this paper appeared. Indeed some 
motivations might be presented, however the main one is simply to gain a 
deeper understanding of quantum channels in a multiparticle setting.
We have put together a modified introduction that reflects that point 
of view, but also addresses the referee's concern, see list of changes.

*********************************************************************
I feel like there is a simple example of the PCE channel: it might be
called the stabilizer state PCE, e.g., N qubit stabilizer state,
erasing one stabilizer generator could give another stabilizer state,
and the remaining Pauli component is obviously 2^{N-1}, satisfying the
2^K components preserving statement. Moreover, in this case erasing
any one stabilizer generator will preserve exactly half the Pauli
components, as generator for all PCE channel?

Since stabilizer states are very useful in quantum computation,
perhaps it is a good practical motivation to study this PCE channel,
as long as they have a solid connection?

Answer:
The connection with stabilizer states might be interesting, given 
that for both stabilizer states and PCEs, Pauli strings play 
a prominent role in their definitions. We have added a paragraph 
at the end of section III pointing out this possible relation.

*********************************************************************
}}}

Answer to of the Second Referee {{{
----------------------------------------------------------------------

*********************************************************************
Could you shed more light on the physical motivation for this study?

Answer:
This paper grew from the sheer curiosity of why, for a single qubit, some
projections are allowed and others not. After some time, the beautiful 
mathematical structure presented in this paper appeared. Indeed some 
motivations might be presented, however the main one is simply to gain a 
deeper understanding of quantum channels in a multiparticle setting.
We have put together a modified introduction that reflects that point 
of view, but also addresses the referee's concern, see list of changes.

*********************************************************************
How do you think the scheme would work for qudits in dimensions of
Hilbert spaces which are not 2^n? Of course, I understand that your
results are pertinent to basis of Pauli tensors, but I am curious to
know your thoughts on the same.

For dimensions $d$ that allow $d+1$ maximally unbiased bases, one can
construct generalized Pauli channels. In those cases, do you think
your results could be extended in some way?

Answer:
This is clearly an interesting question. The generalization is not totally 
direct, and requires some mathematical work.  This is why we did not
present results in this paper, however we will hopefully present a generalization 
for arbitrary dimension. Some care must be taken as some results depend on the
field structure of the set {0,1}, which will not be available for arbitrary 
$d$.

*********************************************************************

}}}

List of changes {{{
* Added to the introduction the following text:
    The main task which we perform in this paper is the identification of the
    conditions which an arbitrary PCE map must satisfy in order to be completely
    positive. The answer turns out to involve a strikingly simple and unexpected
    mathematical structure that is exploited to gain deeper understanding
    on aforementioned channels, as we show in section \ref{sec:vector_spaces}. This
    structure allows for example to describe such channels with a much reduced
    set of parameters (as compared to specifying a list of all erased Pauli
    components) or to define an
    interesting semigroup structure on the set of all PCE channels. 
    Additionally, these channels, are in a sense the simplest possible channels,
    and as such can be used as building bocks of more general channels, for
    instance through convex combination or composition with unitary transformations.
    To summarize succinctly the final result, we show that it is possible
    to assign to every Pauli string a simple PCE channel, obtained by extending
    the system with an ancilla of a single qubit, acting on the combined system by
    a unitary involving the Pauli string and tracing over the ancilla. It then
    follows from our results that all PCE channels arise from such
    channels by composition.
* Between eq. (2) and eq. (3) we replaced 
    Not every such operation  is a quantum channel;
with 
    Not every such map  is a quantum channel;

* In the caption of figure 1, we replaced 
    ... and thus does not correspond to a quantum operation.
with 
    ... and thus does not correspond to a quantum channel.

* After eq. (6), we replaced 
    In addition, for the operation to be trace preserving...
with 
    In addition, for the map to be trace preserving...

* In that same paragraph, we replaced 
    We shall refer to the map  $xx$, with arbitrary values of $\taus$ (only
    restricted by complete positivity) as {\it Pauli diagonal operations}.  
with 
    We shall refer to the map  $xx$, with arbitrary values of $\taus$ (only
    restricted by complete positivity) as {\it Pauli diagonal maps}.

* Before eq. (13) we replaced 
    The matrix representation of the map corresponding to a Pauli diagonal operation is...
with 
    The matrix representation of the map corresponding to a Pauli diagonal map is...

* After eq. (13) we replaced 
    This superoperator is again diagonal in the (multi-qubit) Pauli basis...
with 
    This matrix is again diagonal in the (multi-qubit) Pauli basis...

* We changed the paragraph leading to eq. (23) from
    Let us now work out the nature of the aforementioned connection.  Let the
    components of  $\vbeta$ be such that $\beta_l = \beta \delta_{lk}$ for a
    particular particle index $k$, with $\beta$ such that $\sa_{\alpha_k\beta}=1$.
    Since $\sa_{\alpha 0}=1$ for any $\alpha$, this particular choice of $\vbeta$
    belongs to  $\Phi(\vec\alpha)$, so that \eref{eq:coneccion:dos:indices:A}
    indeed holds. However, for such a choice
    of $\vbeta$ the above relation reduces to 
to
    Let us now work out the nature of the aforementioned connection.  For
    arbitrary $k$ we define a vector $\vbeta_k$ such that
    $\vbeta_k\in\Phi(\vec\alpha)$ as follows: $\vbeta_k$ is zero everywhere except
    for the $k$'th coordinate, which takes a value $\beta$ such that
    $\sa_{\alpha_k\beta}=1$. Since $\sa_{\alpha 0}=1$ for any $\alpha$, this
    particular choice of $\vbeta$ indeed belongs to  $\Phi(\vec\alpha)$, so
    that if \eref{eq:coneccion:dos:indices:A} holds for all
    $\vbeta\in\Phi(\vec\alpha)$, it must hold for that particular $\vbeta_k$,
    which leads to:

* At the end of section III we added the following paragraph:
    Finally, let us point out the following: if we wish to specify a PCE
    channel explicitly, the naive way to proceed would be simply to list all the
    Pauli components which are not erased. This requires in general, however, an
    exponential amount of information: that is, if the system has $N$ qubits, we
    generally require of the order of $2^N$ bits to do this. If, on the other hand,
    we take advantage of the vector space structure of a PCE channel, we only need
    to specify a basis. Since a basis consists of $N$ vectors of length $N$, the
    information required is only of $N^2$ bits, so that we have obtained a very
    substantial improvement by exploiting complete positivity. This is reminiscent
    of a rather similar effect in {\em stabilizer states\/} which can also be
    specified by $N^2$ bits, as opposed to an exponentially large number of basis
    coefficients for arbitrary states.  A stabilizer state is one which is the
    common eigenvector to the eigenvalue 1 of a set of $N$ commuting 
    Pauli strings.  The similarity is highly intriguing, and
    potentially of interest, since stabilizer states are of central importance in
    quantum error correction \cite{Gottesman1997}. 
}}}