### 2022-09-06 documentos/vector_space.tex

\subsection{PCE quantum channels as vector spaces} % {{{
\label{sec:vector_spaces}
% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{

In this subsection we will provide a one-to-one relation between PCE quantum
channels and the subspaces of a discrete vector subspace associated with the
indices $\valpha$ labeling the components of a state; see
\eref{eq:N_qubits_rho}. 
Some established facts about vector spaces will allow
us to derive the main features of PCE quantum channels.

Let us start by recalling that the problem of determining complete positivity
of a PCE map can be recast as determining which coefficients $\tau_\valpha$ 
are mapped via $\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to positive eigenvalues $\lambda_\valpha$, as in \eref{eq:lambda:is:A:tau}. 
% . In components, this would read 
% \begin{equation}
% % \left(
% \sum_\vbeta
% \san_{\valpha\vbeta} \,
% % \right)
% \tau_{\vec \beta}=2^Np_{\vec\alpha}
% \label{eq:A:tau:is:p}
% \end{equation}
% where
% $\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
% \sa_{\alpha_N\beta_N}$. 
Using the fact that $a^{-1}=a/4$, so 
% (so $A = A^{-1}/2^N$) 
\begin{equation}
A^{-1} = \frac{1}{4^N} A,
\end{equation}
we can directly invert \eref{eq:lambda:is:A:tau} to obtain 
\begin{equation}
% \tau_{\vec \alpha}= \san_{\vec\alpha\vec\beta} \, p_\vbeta,
% \tau = 
% \san \lambda =2^N \tau
\sum_\vbeta \san_{\valpha \vbeta} \lambda_\vbeta =2^N \tau_\valpha
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. 
This is a remarkable equation, as it provides a method to diagonalize the
Choi matrix of any Pauli diagonal map.

Other two simple but crucial observations are the following. 
For valid quantum channels it holds 
% \begin{equation}
% % \sum_{
% \end{equation}
% $\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
% $\lambda_\valpha=0$ for all $\valpha$ in an arbitrary subset of 
% multi-indices $\Omega$, 
\begin{equation}
\sum_{\valpha \in \Omega} \lambda_\valpha=0 
\quad \implies 
\quad 
\lambda_\valpha=0, \, \forall \valpha \in \Omega
\label{eq:lambda:is:zero}
\end{equation}
for an arbitrary subset of multi-indices $\Omega$, 
as each member of the sum is greater than or equal to zero, 
due to complete positivity of the underlying channel. 
Finally, setting $\valpha =0$ in \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha)=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
% If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
If we now assume that $\tau_{\vec\alpha}=1$, and calculate the difference between
\eref{eq:sum:alpha:is:2N} and
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$
(which follows from \eref{eq:tau:is:A:p} and $\tau_{\vec\alpha}=1$), one obtains
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three
multi-indices, $\valpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$.

Let us now work out the nature of the aforementioned connection.  For arbitrary
$k$ we define a vector $\vbeta_k$ such that $\vbeta_k\in\Phi(\vec\alpha)$ as
follows: $\vbeta_k$ is zero everywhere except
for the $k$'th coordinate, which takes a value $\beta$ such that  $\sa_{\alpha_k\beta}=1$. Since 
$\sa_{\alpha 0}=1$ for any $\alpha$, this particular choice of $\vbeta$
indeed belongs to  $\Phi(\vec\alpha)$, so that if \eref{eq:coneccion:dos:indices:A} holds for 
all $\vbeta\in\Phi(\vec\alpha)$, it must hold for that particular $\vbeta_k$, which leads to:
% 
% Let us now work out the nature of the aforementioned connection.  Let the
% components of  $\vbeta$ be such that $\beta_l = \beta \delta_{lk}$ for a
% particular particle index $k$, with $\beta$ such that $\sa_{\alpha_k\beta}=1$.
% Since $\sa_{\alpha 0}=1$ for any $\alpha$, this particular choice of $\vbeta$
% belongs to  $\Phi(\vec\alpha)$, so that \eref{eq:coneccion:dos:indices:A}
% indeed holds. However, for such a choice
% of $\vbeta$ the above relation reduces
% to 
\begin{equation}
\sa_{\beta\gamma_k}=\sa_{\beta\gamma_k^\prime}
\label{eq:14}
\end{equation}
for all $\beta$ such that $\sa_{{\alpha_k}\beta}=1$.  One can verify, by
working out the different cases, that \eref{eq:14} is equivalently expressed as 
\begin{equation}
\gamma_k^\prime=\alpha_k\oplus{}\gamma_k
\label{eq:oplus}
\end{equation}
where the $\oplus{}$ denotes the operation of the Klein group; see Table
\ref{tab:1} for a detailed description. 
\par
% }}}
% Espacio vectorial {{{

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$.
We notice that the Klein group described in Table \ref{tab:1} is actually isomorphic
to the two dimensional vector space over the field of two elements $\{0,1\}$.
Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further unravel the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, and say that for quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preserved, and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

\begin{table} % {{{
\begin{center}
\begin{tabular}{| c | c c c c |}
\hline
$\oplus{}$ & 0 & 1& 2 & 3\\
\hline
0 & 0 & 1 & 2 & 3\\
1 & 1 & 0 & 3 & 2\\
2 & 2 & 3 & 0 & 1 \\
3 &3 & 2 & 1 & 0\\
\hline
\end{tabular}
\caption{Definition of the $\oplus{}$ operation, see \eref{eq:14}. Note that
the operation is an abelian group, in fact it corresponds to the 
Klein group, where the neutral element is $0$. This is the reason for choosing
an additive notation for the operation defined in (\ref{eq:14}). }
\label{tab:1}
\end{center}
\end{table} % }}}

From this readily follows an amusing property: the set of all multi-indices
$\vec\gamma$ for which $\tau_{\vec\gamma}=1$, is closed under binary
vector addition, in other words, it forms a vector subspace of the set
of all multi-indices.
A moment's consideration will further show that the above reasoning can be
inverted, that is, that if we set all $\tau_{\vec\gamma}$ equal to one
whenever the $\vec\gamma$ belongs to a given vector subspace of the set of
all indices, then $\tau$ indeed has an image
which has
only positive components. 
In other words, there is a one-to-one correspondence between 
a quantum channel, and a vector subspace of the aforementioned space. 
\par

% }}}
% Generacion de soluciones {{{
With this information, we present a procedure to generate all solutions: we start out
from the solution having $\tau_{\vec0}=1$, with everything else $0$. We may
then successively switch $\tau_{\vec\alpha}$'s
% \janote{agregué el subíndice $\vec\alpha$} 
to $1$ for various values of $\valpha$,
taking care immediately to set equal to one the components of $\tau$
% \janote{puse la $\tau$ como vector y hay otro antes}
that
correspond to values of $\vbeta$ generated by the previously switched values of
$\valpha$ via the operation $\oplus$. Doing so, in an ordered way, allows one
to generate
all PCE quantum channels with a given set of preserved components, without the need
of exploring the exponentially large space of all PCE maps.  \par 
% }}}}

\begin{figure} % {{{
\begin{center}
\includegraphics{two_qubit_channels.pdf}
% Del dcumento francois.pdf pg 8 agarramos:
% e11, e4 y E7
% \begin{tikzpicture}[x=0.55cm, y=0.55cm] % {{{
% \pgfmathsetmacro{\unitstep}{5.2}
% \begin{scope}[shift={(1*\unitstep,0)}] % Depolirizng channel {{{
% \node at (-0.5,0.5) {(a)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \begin{scope}[shift={(2*\unitstep,0)}] % Interseccion {{{
% \node at (-0.5,0.5) {(b)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% % \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \begin{scope}[shift={(3*\unitstep,0)}] % E4 {{{
% \node at (-0.5,0.5) {(c)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% % \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \begin{scope}[shift={(1*\unitstep,-\unitstep)}] % E7 {{{
% \node at (-0.5,0.5) {(d)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% % \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \begin{scope}[shift={(2*\unitstep,-\unitstep)}] % E11 {{{
% \node at (-0.5,0.5) {(e)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% % \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \begin{scope}[shift={(3*\unitstep,-\unitstep)}] % Identity channel {{{
% \node at (-0.5,0.5) {(f)} ;
% \cuadritotikz{}
% \begin{scope}[shift={(0,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(0,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(1,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(2,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,0)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-1)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-2)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \begin{scope}[shift={(3,-3)}] \fill[black] (0,0) rectangle (1,1); \end{scope}
% \end{scope} % }}}
% \end{tikzpicture} % }}}
\end{center}
\caption{
Examples of diagrams for several 2-qubit PCE quantum channels: (a) the totally
depolarizing channel, (b) a PCE channel that preserves four components, the
normalization component, one local component of qubit 2 and two correlations between the 
two qubits.
(c), (d) and (e) show the three generators $\mcG_{(1,0)}$, $\mcG_{(3,2)}$ and
$\mcG_{(2,2)}$, respectively; the combination (overlap of diagrams) of any two
of them yields the channel in (b).
(f) represents the identity map. 
}
\label{fig:examplesPCE}
\end{figure} % }}}
% Los canales tiene 2^K componente {{{

% **********************************
% 
% We can show that all PCE quantum channels preserve $2^K$ components. 
% In
% fact, with the above picture, it is clearly revealed that the vector space has
% dimension $2N$. Moreover, 
% consider any subspace and chose the maximum number of linearly independent
% elements (say $K$).  Any of the  $2^K$ subsets of those elements will yield a
% unique element. Moreover, since $\valpha + \valpha = \vec 0$, these are all
% elements of the vector space. This means that any PCE quantum channel will preserve $2^K$ 
% components, for $K$ an integer. See \fref{fig:examplesPCE} for some examples. 
% \par 
% 
% **********************************
% **********************************

We can show that all PCE quantum channels preserve $2^K$ components. 
First recall that a vector space of dimension $d$ over a field of $q$ elements
has $q^d$ elements~\cite{Roman2008}.
Now $V$ is a vector space on a field of two elements having dimension $2N$. 
% The dimension of the vector space $V$ of all multi-indices is $2N$, and 
We have
seen earlier that
\begin{equation}
W=\left\{
\vec\alpha:\vec\alpha\in V, \tau_{\vec\alpha}=1
\right\}
\label{eq:19}
\end{equation}
is a subspace of $V$. 
As such, $W$ has a given dimension $K$, which means that
$W$ has $2^K$ elements. In other words, a set of indices $\tau_\valpha$ with the property
discussed above can only have $2^K$ elements equal to 1, for a given integer
$K$. \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCE quantum channels exist that preserve  $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}\label{eq:conteo:simmetry}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest a relation  between individual channels that
preserve $K$ and $2N-K$ Pauli components that for the time being
has escaped our efforts to identify. 
\par % }}}

% Stabilizer codes {{{
Finally, let us point out the following: if we wish to specify a PCE
channel explicitly, the naive way to proceed would be simply to list all the
Pauli components which are not erased. This requires in general, however, an
exponential amount of information: that is, if the system has $N$ qubits, we
generally require of the order of $2^N$ bits to do this. If, on the other hand,
we take advantage of the vector space structure of a PCE channel, we only need
to specify a basis. Since a basis consists of $N$ vectors of length $N$, the
information required is only of $N^2$ bits, so that we have obtained a very
substantial improvement by exploiting complete positivity. This is reminiscent
of a rather similar effect in {\em stabilizer states\/} which can also be
specified by $N^2$ bits, as opposed to an exponentially large number of basis
coefficients for arbitrary states.  A stabilizer state is one which is the
common eigenvector to the eigenvalue 1 of a set of $N$ commuting 
Pauli strings.  The similarity is highly intriguing, and
potentially of interest, since stabilizer states are of central importance in
quantum error correction \cite{Gottesman1997}.

% }}}
% }}}